### 2013-11-12 proposal/proposal.tex

\documentclass[]{article}
%\usepackage{verbatim}
%\usepackage[dvips]{graphicx}
%\usepackage{pslatex}
\title{Classifying Protein Secondary Structures through Deep Network Learning}
\author{Stephanie deWet and Adam Vail}

\begin{document}
\maketitle

\section{Goals}
The goal of our project is to classify protein structures through the use of deep networks.
Specifically, we are going to attempt to map each amino acid in the protein to its secondary structure.
Protein classification based upon amino acid sequence is an open problem that is well known in machine learning.
Other learning methods have been applied in the past such as k-nearest neighbor, traditional neural networks (single hidden layer), and support vector machines.
With the recent popularity of deep networks, we hope to discover their level of ability to correctly classify protein structures compared with the more conventional learning algorithms.
If time permits, we hope to use a sequential labeling model such as hidden Markov models or conditional random fields to process the output of the deep network model.

\section{Algorithms}
As mentioned above, we plan to apply deep networks to the problem of protein classification.
We plan on doing this through the use of autoencoders.
Our hope is that the multiple levels of autoencoders will discover comprehensible internal structures that combine to make up the secodary structures.

\section{Experiments}
We plan to format the problem as predicting the secondary structure of an amino acid, given a window of amino acids around it.
We will try solving the problem using both a traditional neural net with back propagation and a deep network with autoencoders.
Our goal is to compare the two methods, and see whether deep networks provide a benefit over the traditional neural networks.

\section{Datasets}
We plan to use the 130 non-redundant proteins selected in \cite{Rost}, because they have already been selected intelligently.
We will use the CATH database to get the amino acid sequence for proteins, because it has a clean and convenient API to get this sequence, and
the Database of Secondary Structure in Proteins (DSSP) to provide the correct secondary structure for our training data.

\begin{thebibliography}{9}

\bibitem{Rost}
Burkhard Rost and Christ Sander,
  `` Prediction of Protein Secondary Structures at Bettter than 70\% Accuracy"
  in \emph{J. Mol. Bio.}, 1993.

\end{thebibliography}

\end{document}

### 2013-11-12 preprocessors/rost/pdb_assemble_data.py
translate the structure of the dssp database into the three classification options presented everywhere else

@@ -48,7 +48,34 @@ class AssembleData(object):
                     dssp = line[colon_index + 1:].strip()
                     proteins[protein].append(dssp)
 
-    def output_proteins(object, proteins):
+    # h -> h
+    # g -> h
+    # i -> h
+    # e -> e
+    # b -> _
+    # t -> _
+    # s -> _
+    def clean_classifications(self, proteins):
+        for protein in proteins:
+            orig_classification = proteins[protein]
+            clean_classification = ""
+            for letter in range(len(orig_classification)):
+                if orig_classification[letter] == 'G' or\
+                        orig_classification[letter] == 'I' or\
+                        orig_classification[letter] == 'H':
+                    clean_classification += 'h'
+                elif orig_classification[letter] == 'E':
+                    clean_classification += 'e'
+                elif orig_classification[letter] == 'B' or\
+                        orig_classification[letter] == 'T' or\
+                        orig_classification[letter] == 'S':
+                    clean_classification += '_'
+            if len(clean_classification) != len(orig_classification):
+                print "Classification translation did not come out right!!"
+            proteins[protein] = clean_classification
+#            return proteins
+
+    def output_proteins(self, proteins):
         output = open("protein_data", "w")
         for protein in proteins:
             output.write("%s %s %s\n" % (protein, proteins[protein][0], proteins[protein][1]))
@@ -68,6 +95,6 @@ if __name__ == '__main__':
     ad = AssembleData(sys.argv[1], sys.argv[2])
     proteins = ad.get_protein_list()
     ad.get_data(proteins)
+    ad.clean_classifications(proteins)
     ad.print_proteins(proteins)
     ad.output_proteins(proteins)
-

### 2013-11-12 ProteinSSClassification/src/dataprocessing/CrossValidation.java
starting to create data processing for cross validation

@@ -1,0 +1,13 @@
+package dataprocessing;
+
+public class CrossValidation {
+	
+	public static ProteinDataSet processData(int degree, String filename){
+		
+		return null;
+	}
+	
+	
+	
+
+}

### 2013-11-12 ProteinSSClassification/src/dataprocessing/Protein.java

@@ -1,0 +1,20 @@
+package dataprocessing;
+
+public class Protein {
+
+	private String sequence;
+	private String secondaryStructure;
+	
+	public Protein(String sequence, String secondaryStructure){
+		this.sequence = sequence;
+		this.secondaryStructure = secondaryStructure;
+	}
+	
+	public String getSequence(){
+		return this.sequence;
+	}
+	
+	public String getSecondaryStructure(){
+		return this.secondaryStructure;
+	}
+}

### 2013-11-12 ProteinSSClassification/src/dataprocessing/ProteinDataSet.java

@@ -1,0 +1,31 @@
+package dataprocessing;
+
+import java.util.ArrayList;
+
+public class ProteinDataSet {
+
+	private ArrayList<Protein> train;
+	private ArrayList<Protein> test;
+	
+	public ProteinDataSet(){
+		this.train = new ArrayList<Protein>();
+		this.test = new ArrayList<Protein>();
+	}
+	
+	public void addProteinToTrain(Protein protein){
+		this.train.add(protein);
+	}
+	
+	public void addProteinToTest(Protein protein){
+		this.test.add(protein);
+	}
+	
+	public ArrayList<Protein> getTrain(){
+		return this.train;
+	}
+	
+	public ArrayList<Protein> getTest(){
+		return this.test;
+	}
+
+}

### 2013-11-13 ProteinSSClassification/src/dataprocessing/CrossValidation.java
all that is left is chopping up the data to get the cross validation

@@ -1,13 +1,61 @@
 package dataprocessing;
 
+import java.io.BufferedReader;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+import java.util.ArrayList;
+
 public class CrossValidation {
 	
-	public static ProteinDataSet processData(int degree, String filename){
+	public static ArrayList<ProteinDataSet> processData(int degree, String filename){		
+		return crossValidation(readData(filename));
+	}
+	
+	private static ArrayList<ProteinDataSet> crossValidation(ArrayList<Protein> proteins){
+		
+		// TODO actually chop up the data to accomplish cross validation
 		
 		return null;
 	}
 	
-	
+	private static ArrayList<Protein> readData(String filename){
+		ArrayList<Protein> proteins = new ArrayList<Protein>();
+		BufferedReader reader = null;
+		
+		try {
+			reader = new BufferedReader(new FileReader(filename));
+			String line = reader.readLine();
+			while(line != null){
+				
+				String[] tokens = line.split("[ ]+");
+				if(tokens.length == 3){
+					proteins.add(new Protein(tokens[0], tokens[1], tokens[2]));
+				}
+				else if(tokens.length == 2){
+					proteins.add(new Protein(tokens[0], tokens[1]));
+				}
+				else {
+					System.out.println("DO NOT KNOW HOW TO HANDLE PROTEIN:\n" + line);
+				}
+				
+				line = reader.readLine();
+			}
+			
+		} catch (FileNotFoundException e) {
+			System.out.println("Could not find file: " + filename);
+		} catch (IOException e) {
+			System.out.println("Could not read from file: " + filename);
+		}
+		
+		try {
+			reader.close();
+		} catch (IOException e) {
+			System.out.println("Could not close input file");
+		}
+		
+		return proteins;
+	}
 	
 
 }

### 2013-11-13 ProteinSSClassification/src/classification/ProteinSSClassification.java
cross validation looks to be working

@@ -1,14 +1,51 @@
 package classification;
 
-public class ProteinSSClassification {
+import java.util.ArrayList;
 
+import dataprocessing.CrossValidation;
+import dataprocessing.Protein;
+import dataprocessing.ProteinDataSet;
+
+public class ProteinSSClassification {
+	
+	public ProteinSSClassification(String filename){
+		ArrayList<ProteinDataSet> data = CrossValidation.processData(crossValidationDegree, filename);
+
+		//printDataSets(data);
+	}
+	
+
+	static int crossValidationDegree = 7;
+	
 	/**
 	 * @param args
 	 */
 	public static void main(String[] args) {
 		// Here is the start of our protein secondary
 		// structure classification project
+		
+		if(args.length != 2){
+			System.out.println("Usage: ./ProteinSSClassification <protein train> <protein test>");
+			System.exit(1);
+		}
+		
+		ProteinSSClassification classification = new ProteinSSClassification(args[0]);
 
+	}
+	
+	public void printDataSets(ArrayList<ProteinDataSet> dataSets){
+		for(ProteinDataSet ds : dataSets){
+			for(Protein p : ds.getTrain()){
+				System.out.println(p.getName());
+			}
+			System.out.println("\n");
+			
+			for(Protein p : ds.getTest()){
+				System.out.println(p.getName());
+			}
+			
+			System.out.println("------------------------------\n");
+		}
 	}
 
 }

### 2013-11-13 ProteinSSClassification/src/dataprocessing/CrossValidation.java

@@ -9,14 +9,61 @@ import java.util.ArrayList;
 public class CrossValidation {
 	
 	public static ArrayList<ProteinDataSet> processData(int degree, String filename){		
-		return crossValidation(readData(filename));
+		return crossValidation(degree, readData(filename));
 	}
 	
-	private static ArrayList<ProteinDataSet> crossValidation(ArrayList<Protein> proteins){
+	private static ArrayList<ProteinDataSet> crossValidation(int degree, ArrayList<Protein> proteins){
 		
-		// TODO actually chop up the data to accomplish cross validation
+		ArrayList<ProteinDataSet> fullDataSet = new ArrayList<ProteinDataSet>();
+		ArrayList<ArrayList<Protein>> chunks = new ArrayList<ArrayList<Protein>>();
+		int group_size;
+		int remainder = proteins.size() % degree;
+		int group_index = 1;
 		
-		return null;
+		if(remainder == 0){
+			group_size = proteins.size() / degree;
+		}
+		else {
+			group_size = proteins.size() / (degree - 1);
+		}
+		
+		// make sure the last group isn't going to be too small
+		if((group_size - remainder) > 0){
+			// There is room to give every group one less so that
+			// the final group will have a closer number to all other groups
+			group_size = group_size - ((group_size - remainder) / degree);
+		}
+		
+		// Chunk the data into separate sets
+		for(int i = 1; i <= proteins.size(); i++){
+			if(chunks.size() < group_index){
+				chunks.add(new ArrayList<Protein>());
+			}
+			
+			chunks.get(group_index - 1).add(proteins.get(i - 1));
+			
+			if(i % group_size == 0){
+				group_index++;
+				if(group_index > degree + 1){
+					System.out.println("Too many chunks: " + group_index);
+				}
+			}
+		}
+		
+		// Create the separate groups for the cross validation
+		for(int i = 0; i < degree; i++){
+			ProteinDataSet currDataSet = new ProteinDataSet();
+			for(int k = 0; k < chunks.size(); k++){
+				if(k == i){
+					currDataSet.addProteinListToTest(chunks.get(k));
+				}
+				else {
+					currDataSet.addProteinListToTrain(chunks.get(k));
+				}
+			}
+			fullDataSet.add(currDataSet);
+		}
+		return fullDataSet;
 	}
 	
 	private static ArrayList<Protein> readData(String filename){

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/AutoencoderStochasticGradient.java
Intermediate point

@@ -1,0 +1,47 @@
+package stochastic_gradient;
+import java.lang.Math;
+
+public class AutoencoderStochasticGradient extends StochasticGradient {
+
+	Double[] network(double x, double w) {
+		return -1;
+	}
+
+	/** Calculates error for a specific output unit.
+	 * @param o actual output of this unit
+	 * @param y expected output for this unit
+	 */
+
+	double error(double o, double y) {
+		// Error = (1/2) * sum_i ( (y_i - o_i)^2 )
+		double error_doubled = 0;
+		for (int i = 0; i < o.length, i++) {
+			error_doubled += (y[i] - o[i]) * (y[i] - o[i]);
+		}
+		return error_doubled * 0.5;
+	}
+
+	// Error for a specific training example
+	double error_subgrad(double o, double y, double x) {
+		if (o.length <= 0 || o.length != y.length) {
+			System.out.println("Error in error subgradient calculations!");
+			return -1;
+		}
+		double error_doubled = 0;
+		for (int i = 0; i < o.length, i++) {
+			error_doubled += (y[i] - o[i]) * (y[i] - o[i]);
+		}
+		return error_doubled * 0.5;
+
+
+
+	}
+
+
+	/** 
+	 * @param x input variables
+	 */
+	public trainWeights(ArrayList<ArrayList<Double>> x) {
+		trainWeights(x.toArray(), x.toArray(), x.length);
+	}
+}

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/count_classed_correctly.m

@@ -1,0 +1,12 @@
+function [ count ] = count_classed_correctly( w, y, Z )
+%Returns the number of elements of y that are classified correctly
+
+% Find all the y_i = -1 where w^T*z_i <= -1
+
+count = numel( find((w.' * Z((y == -1), :).') <= -1) );
+
+% Find all the y_i = 1 where w^T*z_i >= 1
+count = count + numel( find((w.' * Z((y == 1), :).') >= 1) );
+
+end
+

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/func.m

@@ -1,0 +1,15 @@
+function [ value ] = func( m, mu, w, y, Z )
+%Support vector classification function for HW 8
+%   Returns the value of the function at the given point.
+
+
+% calculate the penalty vector.
+% penalty(i) is the penalty of the i'th component.
+penalty = max(ones(m, 1) - y .* (Z * w) , zeros(m, 1));
+
+
+% Calculat the value of the full function.
+value = mu * norm(w) + 1/m * sum(penalty);
+
+end
+

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/hw8.m

@@ -1,0 +1,13 @@
+function hw8( )
+%Main method of homework 8
+%   Does all the experiments
+
+data = load ('sg_homework.mat');
+
+alpha_bar = 1.0e4;
+mu = 1.0e-3;
+
+final_val = stochastic_gradient( mu, alpha_bar, data.Z, data.y );
+
+end
+

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/stochastic_gradient.m

@@ -1,0 +1,50 @@
+function [ func_value ] = stochastic_gradient( mu, alpha_bar, Z, y )
+%Classical stochastic gradient function
+%   This function implements the classical stochastic gradient algorithm,
+%   using a weighting parameter of mu, and a minimum step size of
+%   alpha_bar.  Returns the value of the function at the final w.
+
+[m, n] = size(Z);
+
+% Choose initial w at random from the standard normal distribution
+w_next = randn(n, 1);
+
+val = zeros(m + 1, 1); % Save value of function at each value of k
+%classed = zeros(m + 1, 1); % Save # of y classified correctly for each value of k.
+
+
+% Iterate m times
+for k = 1 : m
+    
+    val(k, 1) = func(m, mu, w_next, y, Z);
+    %classed(k, 1) = count_classed_correctly(w_next,y,Z);
+    
+    w = w_next;
+    
+    % Set i_k to an integer between 1 and m inclusive, chosen uniformly at
+    % random.
+    i_k = randi(m);
+    
+    % Calculate the step size for this iteration
+    alpha_k = min([alpha_bar; 1/(k * mu)]);
+    
+    % Calculate the subgradient for the sample i_k
+    g_i_k = subgrad(mu, w, y(i_k), Z(i_k, :).');
+    
+    % Take a step in the negative g_i_k direction
+    w_next = w - alpha_k * g_i_k;
+
+end
+
+val(m+1, 1) = func(m, mu, w_next, y, Z);
+%classed(m+1, 1) = count_classed_correctly(w_next,y,Z);
+
+subplot(2, 1, 1), plot (val)
+%subplot(2, 1, 2), plot (classed)
+
+% plot how many w meet metric, too.
+
+func_value = func(m, mu, w_next, y, Z);
+
+end
+

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/subgrad.m

@@ -1,0 +1,16 @@
+function [ subgrad_value ] = subgrad( mu, w, y_i_k, z_i_k)
+%Subgradient of support vector classification function for HW 8
+%   Returns the value of a subgradient of the function at the given point.
+
+
+subgrad_value = mu * w;
+
+if (y_i_k * w.' * z_i_k < 1)
+    
+    subgrad_value = subgrad_value - y_i_k * z_i_k;
+    
+end
+
+
+end
+

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/AutoencoderStochasticGradient.java
Make sigmoid gradient unit

@@ -1,47 +1,0 @@
-package stochastic_gradient;
-import java.lang.Math;
-
-public class AutoencoderStochasticGradient extends StochasticGradient {
-
-	Double[] network(double x, double w) {
-		return -1;
-	}
-
-	/** Calculates error for a specific output unit.
-	 * @param o actual output of this unit
-	 * @param y expected output for this unit
-	 */
-
-	double error(double o, double y) {
-		// Error = (1/2) * sum_i ( (y_i - o_i)^2 )
-		double error_doubled = 0;
-		for (int i = 0; i < o.length, i++) {
-			error_doubled += (y[i] - o[i]) * (y[i] - o[i]);
-		}
-		return error_doubled * 0.5;
-	}
-
-	// Error for a specific training example
-	double error_subgrad(double o, double y, double x) {
-		if (o.length <= 0 || o.length != y.length) {
-			System.out.println("Error in error subgradient calculations!");
-			return -1;
-		}
-		double error_doubled = 0;
-		for (int i = 0; i < o.length, i++) {
-			error_doubled += (y[i] - o[i]) * (y[i] - o[i]);
-		}
-		return error_doubled * 0.5;
-
-
-
-	}
-
-
-	/** 
-	 * @param x input variables
-	 */
-	public trainWeights(ArrayList<ArrayList<Double>> x) {
-		trainWeights(x.toArray(), x.toArray(), x.length);
-	}
-}

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/SigmoidOutputStochasticGradient.java

@@ -1,0 +1,45 @@
+package stochastic_gradient;
+import java.lang.Math;
+
+public class SigmoidOutputStochasticGradient extends StochasticGradient {
+
+	Double[] network(double x, double w) {
+		return -1;
+	}
+
+	/** Calculates error for a specific output unit.
+	 * @param o actual output of this unit
+	 * @param y expected output for this unit
+	 * @return error for specific output unit
+	 */
+	double error(double o, double y) {
+		// Error = (1/2) * (y - o)^2
+		return 0.5 * (y - o) * (y - o);
+	}
+
+	/** Calculates error gradient for a specific output unit.
+	 * @param o actual output of this unit
+	 * @param y expected output for this unit
+	 * @return error gradient for specific output unit
+	 */
+	Double[] error_subgrad(double o, double y, Double[] x) {
+		if (x.length <= 0)
+			System.out.println("Error in error subgradient calculations!");
+			return -1;
+		}
+		Double[] grad = new Double(x.length);
+		double prefix = (y - o) * o * (1 - o);
+		for (int i = 0; i < x.length, i++) {
+			grad[i] = prefix * x[i];
+		}
+		return grad;
+	}
+
+	/**  Trains the weights for one specific output unit
+	 * @param x arraylist of training instances, which contains an array of x values
+	 * @param y arraylist of training instances' class labels
+	 */
+	public trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
+		trainWeights(x, y);
+	}
+}

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/SigmoidOutputStochasticGradient.java
fix output

@@ -3,8 +3,21 @@ import java.lang.Math;
 
 public class SigmoidOutputStochasticGradient extends StochasticGradient {
 
-	Double[] network(double x, double w) {
-		return -1;
+	/** Calculates output of specific unit  (assumes no constant inputs!)
+	 * @param x array of inputs to unit
+	 * @param w array of weights of unit
+	 * @return expected output of network unit
+	 */
+	double network(Double[] x, Double[] w) {
+		if (x.length <= 0 || x.length != w.length)
+			System.out.println("Error in network calculations!");
+			return -1;
+		}
+		double net = 0;
+		for (int i = 0; i < x.length; i++) {
+			net += x[i] * w[i];
+		}
+		return 1.0 / (1.0 + Math.exp(-1.0 * net));
 	}
 
 	/** Calculates error for a specific output unit.
@@ -38,8 +51,9 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	/**  Trains the weights for one specific output unit
 	 * @param x arraylist of training instances, which contains an array of x values
 	 * @param y arraylist of training instances' class labels
+	 * @return weights input to unit.
 	 */
-	public trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
-		trainWeights(x, y);
+	Double[] trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
+		return trainWeights(x, y);
 	}
 }

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/StochasticGradient.java
Stochastic Gradient for 1 output unit

@@ -1,0 +1,63 @@
+package stochastic_gradient;
+import java.lang.Math;
+
+public abstract class StochasticGradient {
+	double eta;  // Learning rate  (step size)
+	int m;  // Max iters
+
+	StochasticGradient() {
+		eta = 0.1;
+		m = 100;
+	}
+
+	/** Calculates output of specific unit  (assumes no constant inputs!)
+	 * @param x array of inputs to unit
+	 * @param w array of weights of unit
+	 * @return expected output of network unit
+	 */
+	double abstract network(Double[] x, Double[] w);
+
+	/** Calculates error for a specific output unit.
+	 * @param o actual output of this unit
+	 * @param y expected output for this unit
+	 * @return error for specific output unit
+	 */
+	double abstract error(double o, double y);
+
+	/** Calculates error gradient for a specific output unit.
+	 * @param o actual output of this unit
+	 * @param y expected output for this unit
+	 * @return error gradient for specific output unit
+	 */
+	Double[] abstract error_subgrad(double o, double y, Double[] x);
+
+	/**  Trains the weights for one specific output unit
+	 * @param x arraylist of training instances, which contains an array of x values
+	 * @param y arraylist of training instances' class labels
+	 * @return weights input to unit.
+	 */
+	Double[] trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
+		// TODO: more intelligent stopping criteria
+		// TODO:  Allow weights to be initialized more rationally?
+		if (x.length <= 0) {
+			System.out.println("No training instances!");
+			return null;
+		}
+		// Assume each training instance has same length.
+		Double[] w = new Double(x[0].length);
+		for (int i = 0; i < w.length; i++) {
+			w[i] = Math.random();
+		}
+
+		for (int k = 0; k < m; k++) {
+			for (int d = 0; d < x.length(); d++) {
+				double o = network(x[d].toArray(), w);
+				double err = error(o, y[d]);  // Use this for a better stopping criterion
+				Double[] grad_err = error_subgrad(o, y[d], x[d].toArray());
+				for (int i = 0; i < w.length; i++) {
+					w[i] -= eta * grad_err[i];
+				}
+			}
+		}
+	}
+}

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/SigmoidOutputStochasticGradient.java
Continue work on single output unit
Also delete all those unnecessary matlab files

@@ -8,7 +8,7 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	 * @param w array of weights of unit
 	 * @return expected output of network unit
 	 */
-	double network(Double[] x, Double[] w) {
+	public double network(Double[] x, Double[] w) {
 		if (x.length <= 0 || x.length != w.length)
 			System.out.println("Error in network calculations!");
 			return -1;

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/StochasticGradient.java

@@ -1,5 +1,6 @@
 package stochastic_gradient;
 import java.lang.Math;
+import java.util.ArrayList;
 
 public abstract class StochasticGradient {
 	double eta;  // Learning rate  (step size)
@@ -15,28 +16,28 @@ public abstract class StochasticGradient {
 	 * @param w array of weights of unit
 	 * @return expected output of network unit
 	 */
-	double abstract network(Double[] x, Double[] w);
+	public abstract double network(Double[] x, Double[] w);
 
 	/** Calculates error for a specific output unit.
 	 * @param o actual output of this unit
 	 * @param y expected output for this unit
 	 * @return error for specific output unit
 	 */
-	double abstract error(double o, double y);
+	public abstract double error(double o, double y);
 
 	/** Calculates error gradient for a specific output unit.
 	 * @param o actual output of this unit
 	 * @param y expected output for this unit
 	 * @return error gradient for specific output unit
 	 */
-	Double[] abstract error_subgrad(double o, double y, Double[] x);
+	public abstract Double[] error_subgrad(double o, double y, Double[] x);
 
 	/**  Trains the weights for one specific output unit
 	 * @param x arraylist of training instances, which contains an array of x values
 	 * @param y arraylist of training instances' class labels
 	 * @return weights input to unit.
 	 */
-	Double[] trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
+	Double[] trainWeights(ArrayList<ArrayList<Double>x>, ArrayList<Double> y) {
 		// TODO: more intelligent stopping criteria
 		// TODO:  Allow weights to be initialized more rationally?
 		if (x.length <= 0) {

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/count_classed_correctly.m

@@ -1,12 +1,0 @@
-function [ count ] = count_classed_correctly( w, y, Z )
-%Returns the number of elements of y that are classified correctly
-
-% Find all the y_i = -1 where w^T*z_i <= -1
-
-count = numel( find((w.' * Z((y == -1), :).') <= -1) );
-
-% Find all the y_i = 1 where w^T*z_i >= 1
-count = count + numel( find((w.' * Z((y == 1), :).') >= 1) );
-
-end
-

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/func.m

@@ -1,15 +1,0 @@
-function [ value ] = func( m, mu, w, y, Z )
-%Support vector classification function for HW 8
-%   Returns the value of the function at the given point.
-
-
-% calculate the penalty vector.
-% penalty(i) is the penalty of the i'th component.
-penalty = max(ones(m, 1) - y .* (Z * w) , zeros(m, 1));
-
-
-% Calculat the value of the full function.
-value = mu * norm(w) + 1/m * sum(penalty);
-
-end
-

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/hw8.m

@@ -1,13 +1,0 @@
-function hw8( )
-%Main method of homework 8
-%   Does all the experiments
-
-data = load ('sg_homework.mat');
-
-alpha_bar = 1.0e4;
-mu = 1.0e-3;
-
-final_val = stochastic_gradient( mu, alpha_bar, data.Z, data.y );
-
-end
-

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/stochastic_gradient.m

@@ -1,50 +1,0 @@
-function [ func_value ] = stochastic_gradient( mu, alpha_bar, Z, y )
-%Classical stochastic gradient function
-%   This function implements the classical stochastic gradient algorithm,
-%   using a weighting parameter of mu, and a minimum step size of
-%   alpha_bar.  Returns the value of the function at the final w.
-
-[m, n] = size(Z);
-
-% Choose initial w at random from the standard normal distribution
-w_next = randn(n, 1);
-
-val = zeros(m + 1, 1); % Save value of function at each value of k
-%classed = zeros(m + 1, 1); % Save # of y classified correctly for each value of k.
-
-
-% Iterate m times
-for k = 1 : m
-    
-    val(k, 1) = func(m, mu, w_next, y, Z);
-    %classed(k, 1) = count_classed_correctly(w_next,y,Z);
-    
-    w = w_next;
-    
-    % Set i_k to an integer between 1 and m inclusive, chosen uniformly at
-    % random.
-    i_k = randi(m);
-    
-    % Calculate the step size for this iteration
-    alpha_k = min([alpha_bar; 1/(k * mu)]);
-    
-    % Calculate the subgradient for the sample i_k
-    g_i_k = subgrad(mu, w, y(i_k), Z(i_k, :).');
-    
-    % Take a step in the negative g_i_k direction
-    w_next = w - alpha_k * g_i_k;
-
-end
-
-val(m+1, 1) = func(m, mu, w_next, y, Z);
-%classed(m+1, 1) = count_classed_correctly(w_next,y,Z);
-
-subplot(2, 1, 1), plot (val)
-%subplot(2, 1, 2), plot (classed)
-
-% plot how many w meet metric, too.
-
-func_value = func(m, mu, w_next, y, Z);
-
-end
-

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/subgrad.m

@@ -1,16 +1,0 @@
-function [ subgrad_value ] = subgrad( mu, w, y_i_k, z_i_k)
-%Subgradient of support vector classification function for HW 8
-%   Returns the value of a subgradient of the function at the given point.
-
-
-subgrad_value = mu * w;
-
-if (y_i_k * w.' * z_i_k < 1)
-    
-    subgrad_value = subgrad_value - y_i_k * z_i_k;
-    
-end
-
-
-end
-

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Unit.java
added gradient code to unit

@@ -8,18 +8,22 @@ public class Unit {
 
 	private HashMap<Unit, Double> inputs;
 	double value;
-	private Vector<Unit> outputs;
+	private HashMap<Unit, Double> outputs;
+	private boolean outputUnit = false;
+	private double error;
 	
-	public Unit(){
+	public Unit(boolean outputUnit){
 		inputs = new HashMap<Unit, Double>();
-		outputs = new Vector<Unit>();
+		outputs = new HashMap<Unit, Double>();
+		this.outputUnit = outputUnit;
 	}
 	
-	public Unit(Collection<Unit> inputs){
+	public Unit(Collection<Unit> inputs, boolean outputUnit){
 		addAllInput(inputs);
+		this.outputUnit = outputUnit;
 	}
 	
-	public void calculateValue(){
+	public void calculateOutputValue(){
 		double inputSum = 0;
 		for(Unit in : inputs.keySet()){
 			inputSum += in.getValue() * inputs.get(in);
@@ -53,7 +57,7 @@ public class Unit {
 		return inputs;
 	}
 	
-	public Vector<Unit> getOutputs(){
+	public HashMap<Unit, Double> getOutputs(){
 		return outputs;
 	}
 	
@@ -61,4 +65,45 @@ public class Unit {
 		return value;
 	}
 	
+	public double getError() {
+		return error;
+	}
+	
+	public double getWeight(Unit u){
+		return inputs.get(u);
+	}
+	
+	private void calculateOutputUnitError(double y){
+		error = (y - value) * value * (1 - value);
+	}
+	
+	private void calculateHiddenUnitError(){
+		double outputError = 0;
+		for(Unit u : outputs.keySet()){
+			outputError += u.getError() * outputs.get(u);
+			outputs.put(u, u.getWeight(this));
+		}
+		
+		error = value * (1 - value) * outputError;
+	}
+	
+	private void changeWeights(double learningRate) {
+		for(Unit u : inputs.keySet()){
+			double weight = inputs.get(u) + (error * u.getValue() * learningRate);
+			inputs.put(u, weight);
+		}
+	}
+	
+	public void trainWeights(double y, double learningRate){
+		if(outputUnit){
+			// train weights as an output unit
+			calculateOutputUnitError(y);
+		}
+		else {
+			// train weights as a hidden unit
+			calculateHiddenUnitError();
+		}
+		
+		changeWeights(learningRate);
+	}
 }

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/SigmoidOutputStochasticGradient.java

@@ -1,5 +1,6 @@
 package stochastic_gradient;
-import java.lang.Math;
+
+import java.util.ArrayList;
 
 public class SigmoidOutputStochasticGradient extends StochasticGradient {
 
@@ -9,7 +10,7 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	 * @return expected output of network unit
 	 */
 	public double network(Double[] x, Double[] w) {
-		if (x.length <= 0 || x.length != w.length)
+		if (x.length <= 0 || x.length != w.length) {
 			System.out.println("Error in network calculations!");
 			return -1;
 		}
@@ -25,7 +26,7 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	 * @param y expected output for this unit
 	 * @return error for specific output unit
 	 */
-	double error(double o, double y) {
+	public double error(double o, double y) {
 		// Error = (1/2) * (y - o)^2
 		return 0.5 * (y - o) * (y - o);
 	}
@@ -35,14 +36,15 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	 * @param y expected output for this unit
 	 * @return error gradient for specific output unit
 	 */
-	Double[] error_subgrad(double o, double y, Double[] x) {
-		if (x.length <= 0)
+	public Double[] error_subgrad(double o, double y, Double[] x) {
+		if (x.length <= 0) {
 			System.out.println("Error in error subgradient calculations!");
-			return -1;
+			return null;
 		}
-		Double[] grad = new Double(x.length);
+		Double[] grad = new Double[x.length];
+		
 		double prefix = (y - o) * o * (1 - o);
-		for (int i = 0; i < x.length, i++) {
+		for (int i = 0; i < x.length; i++) {
 			grad[i] = prefix * x[i];
 		}
 		return grad;
@@ -53,7 +55,7 @@ public class SigmoidOutputStochasticGradient extends StochasticGradient {
 	 * @param y arraylist of training instances' class labels
 	 * @return weights input to unit.
 	 */
-	Double[] trainWeights(ArrayList<ArrayList<Double> x>, ArrayList<Double> y) {
+	public Double[] trainWeights(ArrayList<ArrayList<Double>> x, ArrayList<Double> y) {
 		return trainWeights(x, y);
 	}
 }

### 2013-11-14 ProteinSSClassification/src/stochastic_gradient/StochasticGradient.java

@@ -37,28 +37,29 @@ public abstract class StochasticGradient {
 	 * @param y arraylist of training instances' class labels
 	 * @return weights input to unit.
 	 */
-	Double[] trainWeights(ArrayList<ArrayList<Double>x>, ArrayList<Double> y) {
+	public Double[] trainWeights(ArrayList<ArrayList<Double>> x, ArrayList<Double> y) {
 		// TODO: more intelligent stopping criteria
 		// TODO:  Allow weights to be initialized more rationally?
-		if (x.length <= 0) {
+		if (x.size() <= 0) {
 			System.out.println("No training instances!");
 			return null;
 		}
 		// Assume each training instance has same length.
-		Double[] w = new Double(x[0].length);
+		Double[] w = new Double[x.get(0).size()];
 		for (int i = 0; i < w.length; i++) {
 			w[i] = Math.random();
 		}
 
 		for (int k = 0; k < m; k++) {
-			for (int d = 0; d < x.length(); d++) {
-				double o = network(x[d].toArray(), w);
-				double err = error(o, y[d]);  // Use this for a better stopping criterion
-				Double[] grad_err = error_subgrad(o, y[d], x[d].toArray());
+			for (int d = 0; d < x.size(); d++) {
+				double o = network((Double[])x.get(d).toArray(), w);
+				double err = error(o, y.get(d));  // Use this for a better stopping criterion
+				Double[] grad_err = error_subgrad(o, y.get(d), (Double[])x.get(d).toArray());
 				for (int i = 0; i < w.length; i++) {
 					w[i] -= eta * grad_err[i];
 				}
 			}
 		}
+		return w;
 	}
 }

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Unit.java
added function to set the unit value for input values

@@ -53,6 +53,10 @@ public class Unit {
 		}
 	}
 	
+	public void setValue(double value){
+		this.value = value;
+	}
+	
 	public HashMap<Unit, Double> getInputs(){
 		return inputs;
 	}

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Controller.java
Start building a Unit Controller

@@ -1,0 +1,58 @@
+package neuralnetwork;
+
+import java.util.Collection;
+import java.util.HashMap;
+import java.util.Vector;
+
+public class Controller {
+
+	private Collection<Collection<Unit>> allUnits;
+
+	// Input is size of each layer of units.  Ex: 17 5 3
+	public Controller(ArrayList<int> size){
+		for(int i = 0; i < size.size(); i++) {
+			Collection<Unit> layer = new Collection<Units>();
+			for (int j = 0; j < size.get(i); j++) { 
+				if (i < size.size() - 1) {
+					layer.add(new Unit(false));
+				}
+				else {
+					layer.add(new Unit(true));
+				}
+			}
+			allUnits.add(layer);
+		}
+
+		// Need to set up connections
+	}
+
+	// Train network on a piece of data
+	public void trainOnInstance(ArrayList<double> x, y) {
+		if (x.size() != inputs.size()) {
+			// Error
+			System.out.println("Error in training instances");
+			return;
+		}
+		// Set the values of each element of the network, starting with first
+		// layer of network and working forward
+		for (int i = 0; i < x.size(); i++) {
+			inputs.get(i).setValue(x.get(i));
+		}
+
+		// Train weights, starting with back row of network and working backward.
+		trainWeights(y,learningRate);
+	}
+
+	private void setNetworkValues() {
+		// Initialize Input layer
+		for (int i = 0; i < x.size(); i++) {
+			allUnits.get(0).get(i).setValue(x.get(i));
+		}
+		// Initialize other layers
+		for (Collection<Unit> layer : allUnits) {
+			for (Unit unit : layer) {
+			}
+		}
+
+	}
+}

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Unit.java

@@ -17,9 +17,17 @@ public class Unit {
 		outputs = new HashMap<Unit, Double>();
 		this.outputUnit = outputUnit;
 	}
-	
+
+	// Use this constructor for output units
 	public Unit(Collection<Unit> inputs, boolean outputUnit){
 		addAllInput(inputs);
+		this.outputUnit = outputUnit;
+	}
+	
+	// Use this constructor for hidden units
+	public Unit(Collection<Unit> inputs, Collection<Unit> outputs, boolean outputUnit){
+		addAllInput(inputs);
+		addAllOutput(outputs);
 		this.outputUnit = outputUnit;
 	}
 	
@@ -50,6 +58,24 @@ public class Unit {
 			// get a random double between -1 and 1
 			double randomWeight = (Math.random() * 2) - 1;
 			this.inputs.put(u, randomWeight);
+		}
+	}
+	
+	public void addOutput(Unit output, double weight){
+		outputs.put(output, weight);
+	}
+	
+	public void addOutput(Unit output){
+		// get a random double between -1 and 1
+		double randomWeight = (Math.random() * 2) - 1;
+		this.outputs.put(output, randomWeight);
+	}
+	
+	public void addAllOutput(Collection<Unit> outputs){
+		for(Unit u : outputs){
+			// get a random double between -1 and 1
+			double randomWeight = (Math.random() * 2) - 1;
+			this.outputs.put(u, randomWeight);
 		}
 	}
 	

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Controller.java
initialized the network, need to feed forward and do back propagation

@@ -1,33 +1,86 @@
 package neuralnetwork;
 
+import java.util.ArrayList;
 import java.util.Collection;
-import java.util.HashMap;
-import java.util.Vector;
+
+import dataprocessing.ProteinDataSet;
 
 public class Controller {
 
-	private Collection<Collection<Unit>> allUnits;
+	// handle to our network
+	private ArrayList<ArrayList<Unit>> allUnits = new ArrayList<ArrayList<Unit>>();
+	
+	private ArrayList<InputUnit> inputs;
+	int windowSize;
+	final int aminoAcidLibrary = 20;
 
 	// Input is size of each layer of units.  Ex: 17 5 3
-	public Controller(ArrayList<int> size){
-		for(int i = 0; i < size.size(); i++) {
-			Collection<Unit> layer = new Collection<Units>();
-			for (int j = 0; j < size.get(i); j++) { 
-				if (i < size.size() - 1) {
-					layer.add(new Unit(false));
+	public Controller(ArrayList<Integer> layerSize, int windowSize){
+		this.windowSize = windowSize;
+		
+		initializeNeuralNetwork(layerSize);
+		
+	}
+	
+	public ArrayList<ArrayList<Unit>> getNetwork(){
+		return allUnits;
+	}
+	
+	public void initializeNeuralNetwork(ArrayList<Integer> layerSize){
+		// initialize the units array list
+		for(int i = 0; i < layerSize.size(); i++){
+			allUnits.add(new ArrayList<Unit>());
+		}
+		
+		createInputs();
+		
+		for(int i = 0; i < layerSize.size(); i++) {
+			for (int j = 0; j < layerSize.get(i); j++) { 
+				if (i < layerSize.size() - 1) {
+					allUnits.get(i).add(new Unit(false));
 				}
 				else {
-					layer.add(new Unit(true));
+					allUnits.get(i).add(new Unit(true));
 				}
 			}
-			allUnits.add(layer);
 		}
-
-		// Need to set up connections
+		
+		connectGraph();
+	}
+	
+	private void connectGraph(){
+		
+		// connect all the inputs.
+		for(int i = 0; i < allUnits.size() - 1; i++){
+			for(Unit u : allUnits.get(i+1)){
+				u.addAllInput(allUnits.get(i));
+			}
+		}
+		
+		// set the outputs
+		for(int i = allUnits.size() - 1; i > 0; i--){
+			for(Unit out : allUnits.get(i)){
[…]

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/InputUnit.java

@@ -1,0 +1,38 @@
+package neuralnetwork;
+
+import java.util.ArrayList;
+
+public class InputUnit {
+
+	ArrayList<Unit> subUnits = new ArrayList<Unit>();
+	
+	public InputUnit(int quantity){
+		
+		for(int i = 0; i < quantity; i++){
+			subUnits.add(new Unit(false));
+		}
+		
+	}
+	
+	public ArrayList<Unit> getInputSubUnits(){
+		return subUnits;
+	}
+	
+	public void setInputUnit(char aminoAcid){
+		
+		int target = ((int)aminoAcid) - 65;
+		if(target > subUnits.size()){
+			System.out.println("Bad amino acid: " + aminoAcid);
+		}
+		
+		for(int i = 0; i < subUnits.size(); i++){
+			if(i == target){
+				subUnits.get(i).setValue(1);
+			}
+			else {
+				subUnits.get(i).setValue(0);
+			}
+		}
+	}
+	
+}

### 2013-11-14 ProteinSSClassification/src/neuralnetwork/Unit.java

@@ -2,19 +2,17 @@ package neuralnetwork;
 
 import java.util.Collection;
 import java.util.HashMap;
-import java.util.Vector;
+import java.util.Map;
 
 public class Unit {
 
-	private HashMap<Unit, Double> inputs;
+	private HashMap<Unit, Double> inputs = new HashMap<Unit, Double>();
 	double value;
-	private HashMap<Unit, Double> outputs;
+	private HashMap<Unit, Double> outputs = new HashMap<Unit, Double>();
 	private boolean outputUnit = false;
 	private double error;
 	
 	public Unit(boolean outputUnit){
-		inputs = new HashMap<Unit, Double>();
-		outputs = new HashMap<Unit, Double>();
 		this.outputUnit = outputUnit;
 	}
 
@@ -25,7 +23,7 @@ public class Unit {
 	}
 	
 	// Use this constructor for hidden units
-	public Unit(Collection<Unit> inputs, Collection<Unit> outputs, boolean outputUnit){
+	public Unit(Collection<Unit> inputs, Map<Unit, Double> outputs, boolean outputUnit){
 		addAllInput(inputs);
 		addAllOutput(outputs);
 		this.outputUnit = outputUnit;
@@ -71,11 +69,10 @@ public class Unit {
 		this.outputs.put(output, randomWeight);
 	}
 	
-	public void addAllOutput(Collection<Unit> outputs){
-		for(Unit u : outputs){
+	public void addAllOutput(Map<Unit, Double> outputs){
+		for(Unit u : outputs.keySet()){
 			// get a random double between -1 and 1
-			double randomWeight = (Math.random() * 2) - 1;
-			this.outputs.put(u, randomWeight);
+			this.outputs.put(u, outputs.get(u));
 		}
 	}
 	

### 2013-11-15 ProteinSSClassification/src/neuralnetwork/Controller.java
lower layer neural net written, untested

@@ -13,6 +13,7 @@ public class Controller {
 	private ArrayList<InputUnit> inputs;
 	int windowSize;
 	final int aminoAcidLibrary = 20;
+	final double LEARNING_RATE = .1;
 
 	// Input is size of each layer of units.  Ex: 17 5 3
 	public Controller(ArrayList<Integer> layerSize, int windowSize){
@@ -50,7 +51,7 @@ public class Controller {
 	
 	private void connectGraph(){
 		
-		// connect all the inputs.
+		// connect all the inputs. Yields initial wedge weights
 		for(int i = 0; i < allUnits.size() - 1; i++){
 			for(Unit u : allUnits.get(i+1)){
 				u.addAllInput(allUnits.get(i));
@@ -80,33 +81,76 @@ public class Controller {
 	// Train network on a piece of data
 	public void trainOnInstance(String sequence, STRUCTURE structure) {
 		// set inputs to correct value
+		setInputs(sequence);
+		
 		// loop through layers
-		if (x.size() != inputs.size()) {
-			// Error
-			System.out.println("Error in training instances");
-			return;
-		}
-		// Set the values of each element of the network, starting with first
-		// layer of network and working forward
-		for (int i = 0; i < x.size(); i++) {
-			inputs.get(i).setValue(x.get(i));
-		}
-
-		// Train weights, starting with back row of network and working backward.
-		trainWeights(y,learningRate);
+		feedForward();
+		backPropagate(structure);
 	}
-
-	private void setNetworkValues() {
-		// Initialize Input layer
-		for (int i = 0; i < x.size(); i++) {
-			allUnits.get(0).get(i).setValue(x.get(i));
+	
+	private void setInputs(String sequence){
+		if(sequence.length() != inputs.size()){
+			System.out.println("Sequence length and number of inputs are not equal!");
+			System.exit(1);
 		}
-		// Initialize other layers
-		for (Collection<Unit> layer : allUnits) {
-			for (Unit unit : layer) {
+		
+		for(int i = 0; i < sequence.length(); i++){
+			inputs.get(i).setInputUnit(sequence.charAt(i));
+		}
+	}
+	
+	private void feedForward(){
+		for(int i = 1; i < allUnits.size(); i++){
+			for(Unit u : allUnits.get(i)){
[…]

### 2013-11-15 ProteinSSClassification/src/neuralnetwork/Unit.java

@@ -4,6 +4,8 @@ import java.util.Collection;
 import java.util.HashMap;
 import java.util.Map;
 
+import neuralnetwork.Controller.STRUCTURE;
+
 public class Unit {
 
 	private HashMap<Unit, Double> inputs = new HashMap<Unit, Double>();
@@ -100,7 +102,7 @@ public class Unit {
 		return inputs.get(u);
 	}
 	
-	private void calculateOutputUnitError(double y){
+	private void calculateOutputUnitError(double y){		
 		error = (y - value) * value * (1 - value);
 	}
 	
@@ -121,7 +123,7 @@ public class Unit {
 		}
 	}
 	
-	public void trainWeights(double y, double learningRate){
+	public void trainWeights(Double y, double learningRate){
 		if(outputUnit){
 			// train weights as an output unit
 			calculateOutputUnitError(y);

### 2013-11-15 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java
coming together, both versions of learning. One for AE and another for NN

@@ -1,52 +1,52 @@
 package neuralnetwork;
 
 import java.util.ArrayList;
-import java.util.Collection;
 
-import dataprocessing.ProteinDataSet;
-
-public class Controller {
+public class NeuralNetworkController {
 
 	// handle to our network
 	private ArrayList<ArrayList<Unit>> allUnits = new ArrayList<ArrayList<Unit>>();
 	
-	private ArrayList<InputUnit> inputs;
-	int windowSize;
 	final int aminoAcidLibrary = 20;
 	final double LEARNING_RATE = .1;
+	final int OUTPUT_LAYER_SIZE = 3;
 
 	// Input is size of each layer of units.  Ex: 17 5 3
-	public Controller(ArrayList<Integer> layerSize, int windowSize){
-		this.windowSize = windowSize;
-		
-		initializeNeuralNetwork(layerSize);
-		
-	}
-	
-	public ArrayList<ArrayList<Unit>> getNetwork(){
-		return allUnits;
-	}
-	
-	public void initializeNeuralNetwork(ArrayList<Integer> layerSize){
-		// initialize the units array list
-		for(int i = 0; i < layerSize.size(); i++){
-			allUnits.add(new ArrayList<Unit>());
-		}
-		
-		createInputs();
-		
-		for(int i = 0; i < layerSize.size(); i++) {
-			for (int j = 0; j < layerSize.get(i); j++) { 
-				if (i < layerSize.size() - 1) {
-					allUnits.get(i).add(new Unit(false));
-				}
-				else {
-					allUnits.get(i).add(new Unit(true));
-				}
-			}
-		}
-		
+	public NeuralNetworkController(int inputSize, int hiddenLayerSize){		
+		createLayer(inputSize, false);
+		createLayer(hiddenLayerSize, false);
+		createLayer(inputSize, true);
 		connectGraph();
+	}
+	
+	public NeuralNetworkController(int inputSize){
+		createLayer(inputSize, false);
+		createLayer(OUTPUT_LAYER_SIZE, true);
+		connectGraph();
+	}
+	
+	public ArrayList<Unit> autoencoderLearn(ArrayList<ArrayList<Double>> inputValues){
+		for(ArrayList<Double> inst : inputValues){
+			setInputs(inst);
+			feedForward();
+			backPropagateAE(inst);
+		}
+		
+		// only return the hidden unit layer
+		return allUnits.get(1);
+	}
+	
[…]

### 2013-11-15 ProteinSSClassification/src/neuralnetwork/Unit.java

@@ -4,8 +4,6 @@ import java.util.Collection;
 import java.util.HashMap;
 import java.util.Map;
 
-import neuralnetwork.Controller.STRUCTURE;
-
 public class Unit {
 
 	private HashMap<Unit, Double> inputs = new HashMap<Unit, Double>();
@@ -21,6 +19,11 @@ public class Unit {
 	// Use this constructor for output units
 	public Unit(Collection<Unit> inputs, boolean outputUnit){
 		addAllInput(inputs);
+		this.outputUnit = outputUnit;
+	}
+	
+	public Unit(double value, boolean outputUnit){
+		this.value = value;
 		this.outputUnit = outputUnit;
 	}
 	

### 2013-12-14 ProteinSSClassification/src/classification/ProteinSSClassification.java
Add optional decayBy parameter
This is the fraction of the original amount that you want to decay by
for each layer.
For example, with 45 nodes in the original layer, and decayBy = 1/3,
the layers would have sizes:  45, 30, 15.

@@ -15,23 +15,24 @@ import dataprocessing.ProteinDataSet;
 
 public class ProteinSSClassification {
 	
-	static int crossValidationDegree = 7;
+	static int crossValidationDegree = 5;
 	BufferedWriter outputFile = null;
 	
 	public ProteinSSClassification(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,
-			int hiddenLayerSize, int iterations, String outputDir, boolean baseline){
+			int hiddenLayerSize, int iterations, String outputDir, boolean baseline, double decayBy){
 
 		if(baseline) {
 			runBaseline(trainFilename, testFilename, hiddenLayerSize, windowSize, outputDir);
 		}
 		else {
-			runDeepNetwork(trainFilename, testFilename, windowSize, numHiddenLayers, hiddenLayerSize, iterations, outputDir);
+			runDeepNetwork(trainFilename, testFilename, windowSize, numHiddenLayers, hiddenLayerSize, 
+					iterations, outputDir, decayBy);
 		}
 		 
 	}
 	
 	public void runDeepNetwork(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,
-			int hiddenLayerSize, int iterations, String outputDir) {
+			int hiddenLayerSize, int iterations, String outputDir, double decayBy) {
 		ArrayList<ProteinDataSet> data;
 		
 		Calendar c = Calendar.getInstance();
@@ -80,8 +81,9 @@ public class ProteinSSClassification {
 			
 			AutoencoderController controller = new AutoencoderController(d, windowSize, outputFile);
 			controller.learnInitialLayer(hiddenLayerSize, iterations);
+			int decayAmt = (int)(hiddenLayerSize * decayBy);
 			for(int i = 0; i < numHiddenLayers; i++) {
-				controller.learnHiddenLayer(hiddenLayerSize, iterations);
+				controller.learnHiddenLayer(hiddenLayerSize - i * decayAmt, iterations);
 			}
 			controller.learnOutputLayer(iterations);			
 
@@ -152,7 +154,7 @@ public class ProteinSSClassification {
 		// Here is the start of our protein secondary
 		// structure classification project
 		
-		if(args.length != 8){
+		if(!(args.length == 8 || args.length == 9)){
 			System.out.println("Usage: ./ProteinSSClassification <protein train> <protein test> " + 
[…]

### 2013-12-14 experiments/runExperiments.sh
numLayers tests

@@ -17,7 +17,7 @@ for i in 1 2 3
 do
     outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
     echo "Running Deep Network: hidden layer size 15, layer number -> $i" >> "${outputDirectory}/deep.log"
-    java -jar deepnetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
+    java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
     echo "Done" >> "${outputDirectory}/deep.log"
 done
 
@@ -26,6 +26,6 @@ for i in 1 2 3
 do
     outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
     echo "Running Deep Network: hidden layer size 26, layer number -> $i" >> "${outputDirectory}/deep.log"
-    java -jar deepnetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
+    java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
     echo "Done" >> "${outputDirectory}/deep.log"
 done

### 2013-12-14 ProteinSSClassification/src/classification/ProteinSSClassification.java
Protein Classification

@@ -15,7 +15,7 @@ import dataprocessing.ProteinDataSet;
 
 public class ProteinSSClassification {
 	
-	static int crossValidationDegree = 5;
+	static int crossValidationDegree = 7;
 	BufferedWriter outputFile = null;
 	
 	public ProteinSSClassification(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,

### 2013-12-15 ProteinSSClassification/src/classification/ProteinSSClassification.java
added output connection to all inputs as a flag

@@ -15,23 +15,23 @@ import dataprocessing.ProteinDataSet;
 
 public class ProteinSSClassification {
 	
-	static int crossValidationDegree = 7;
+	static int crossValidationDegree = 5;
 	BufferedWriter outputFile = null;
 	
 	public ProteinSSClassification(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,
-			int hiddenLayerSize, int iterations, String outputDir, boolean baseline){
+			int hiddenLayerSize, int iterations, String outputDir, boolean baseline, boolean traditionalOutput){
 
 		if(baseline) {
 			runBaseline(trainFilename, testFilename, hiddenLayerSize, windowSize, outputDir);
 		}
 		else {
-			runDeepNetwork(trainFilename, testFilename, windowSize, numHiddenLayers, hiddenLayerSize, iterations, outputDir);
+			runDeepNetwork(trainFilename, testFilename, windowSize, numHiddenLayers, hiddenLayerSize, iterations, outputDir, traditionalOutput);
 		}
 		 
 	}
 	
 	public void runDeepNetwork(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,
-			int hiddenLayerSize, int iterations, String outputDir) {
+			int hiddenLayerSize, int iterations, String outputDir, boolean traditionalOutput) {
 		ArrayList<ProteinDataSet> data;
 		
 		Calendar c = Calendar.getInstance();
@@ -83,7 +83,12 @@ public class ProteinSSClassification {
 			for(int i = 0; i < numHiddenLayers; i++) {
 				controller.learnHiddenLayer(hiddenLayerSize, iterations);
 			}
-			controller.learnOutputLayer(iterations);			
+			if(traditionalOutput) {
+				controller.learnOutputLayer(iterations);
+			}
+			else {
+				controller.learnConnectedOutputLayer(iterations);
+			}
 
 			controller.runTestSet();
 			iter++;
@@ -152,9 +157,9 @@ public class ProteinSSClassification {
 		// Here is the start of our protein secondary
 		// structure classification project
 		
-		if(args.length != 8){
+		if(args.length != 9){
 			System.out.println("Usage: ./ProteinSSClassification <protein train> <protein test> " + 
-						"<window size> <number of hidden layers> <hidden layer size> <iterations> <output directory> <baseline>");
[…]

### 2013-12-15 ProteinSSClassification/src/neuralnetwork/AutoencoderController.java

@@ -238,6 +238,37 @@ public class AutoencoderController {
 		
 	}
 	
+	public void learnConnectedOutputLayer(int iterations) {
+		
+		ArrayList<ArrayList<Unit>> outputLayer = null;
+		
+		// get all units to make a flat input layer for the output layer
+		ArrayList<Unit> flatInputLayer = new ArrayList<Unit>();
+		for(int i = 0 ; i < network.size(); i++) {
+			for(int j = 0; j < network.get(i).size(); j++) {
+				flatInputLayer.add(network.get(i).get(j));
+			}
+		}
+		
+		NeuralNetworkController layerController = new 
+				NeuralNetworkController(flatInputLayer, outputFile);
+		layerController.setMaxIterations(iterations);
+		
+		// Feed everything forward and then hand off the output
+		for(int i = 0; i < iterations; i++) {
+			for (int j = 0; j < processedData.size(); j++) {
+				// Feed this piece of data through the network
+				// Now all values are set
+				feedForward(processedData.get(j));
+				// Do back propagation to learn weights to all nodes
+				outputLayer = layerController.learnConnectedOutput(structures.get(j));
+			}
+		}
+		if(outputLayer != null) {
+			network.add(outputLayer.get(1));
+		}
+	}
+	
 	private STRUCTURE convertOutputsToStructure(){
 		double highestOutput = Double.NEGATIVE_INFINITY;
 		int winner = -1;

### 2013-12-15 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java

@@ -183,9 +183,21 @@ public class NeuralNetworkController {
 		return allUnits;
 	}
 	
+	public ArrayList<ArrayList<Unit>> learnConnectedOutput(STRUCTURE output) {
+		
+		// feed forward to the output nodes
+		feedForward();
+		
+		// back propagate to change weights
+		backPropagateNN(output);
+
+		
+		return allUnits;
+	}
+	
 	private void connectGraph(){
 		
-		// connect all the inputs. Yields initial wedge weights
+		// connect all the inputs. Yields initial edge weights
 		for(int i = 0; i < allUnits.size() - 1; i++){
 			for(Unit u : allUnits.get(i+1)){
 				u.addAllInput(allUnits.get(i));

### 2013-12-15 ProteinSSClassification/src/classification/ProteinSSClassification.java
cleaned up handing in parameters, numbers got messed up on merge, added the initial ensemble code

@@ -161,7 +161,7 @@ public class ProteinSSClassification {
 		// structure classification project
 		
 
-		if(!(args.length == 8 || args.length == 9)){
+		if(!(args.length == 11)){
 			System.out.println("Usage: ./ProteinSSClassification <protein train> <protein test> " + 
 						"<window size> <number of hidden layers> <hidden layer size> <iterations> <output directory> <baseline> <traditional output>");
 			System.exit(1);
@@ -185,7 +185,12 @@ public class ProteinSSClassification {
 		}
 		double decayBy = 1.0;
 		if (args.length == 9) {
-			decayBy = Double.parseDouble(args[8]);
+			decayBy = Double.parseDouble(args[9]);
+		}
+		
+		boolean ensemble = false;
+		if(args[10].equalsIgnoreCase("yes") || args[10].equalsIgnoreCase("y")) {
+			ensemble = true;
 		}
 		
 		if(testFilename.equalsIgnoreCase("none")) {

### 2013-12-15 ProteinSSClassification/src/ensemble/NeuralNetEnsemble.java

@@ -1,0 +1,127 @@
+package ensemble;
+
+import java.util.ArrayList;
+
+import baseline.BaselineNeuralNetwork;
+import baseline.BaselineNeuralNetwork.STRUCTURE;
+import dataprocessing.CrossValidation;
+import dataprocessing.Protein;
+import dataprocessing.ProteinDataSet;
+
+public class NeuralNetEnsemble {
+	
+	String trainName;
+	String testName;
+	int hiddenLayerSize;
+	int windowSize;
+	int ensembleSize = 3;
+	ArrayList<ProteinDataSet> data = new ArrayList<ProteinDataSet>();
+	ArrayList<BaselineNeuralNetwork> networks = new ArrayList<BaselineNeuralNetwork>();
+	ArrayList<Protein> testSet;
+	
+	public NeuralNetEnsemble(String trainName, String testName, int hiddenLayerSize, int windowSize, int ensembleSize) {
+		this.trainName = trainName;
+		this.testName = testName;
+		this.hiddenLayerSize = hiddenLayerSize;
+		this.windowSize = windowSize;
+		this.ensembleSize = ensembleSize;
+		testSet = CrossValidation.readData(testName);
+		
+		bagProteins();
+		createNeuralNetworks();
+		
+	}
+	
+	public void classifyTestInstances() {
+		double correct = 0;
+		double total = 0;
+		for(Protein protein : testSet) {
+			ArrayList<ArrayList<Double>> processedTestData = new ArrayList<ArrayList<Double>>();
+			ArrayList<STRUCTURE> testStructures = new ArrayList<STRUCTURE>();
+			processedTestData.addAll(networks.get(0).convertProteinToDoubles(protein));
+			testStructures.addAll(networks.get(0).convertProteinStructure(protein));
+			
+			for(int i = 0; i < processedTestData.size(); i++) {
+				ArrayList<STRUCTURE> outputs = new ArrayList<STRUCTURE>();
+				for(BaselineNeuralNetwork n : networks) {
+					outputs.add(n.classifySingleInstance(processedTestData.get(i)));
+				}
+				// Determine who was the winner
+				STRUCTURE winner = findWinner(outputs);
+				System.out.println("Prediction: " + winner + " Actual: " + testStructures.get(i));
+				total++;
+				if(testStructures.get(i) == winner) {
+					correct++;
+				}
+			}
+		}
[…]

### 2013-12-15 ProteinSSClassification/src/classification/ProteinSSClassification.java
ensemble worked into main code

@@ -1,8 +1,6 @@
 package classification;
 
 import java.io.BufferedWriter;
-import java.io.File;
-import java.io.FileWriter;
 import java.io.IOException;
 import java.util.ArrayList;
 import java.util.Calendar;
@@ -12,6 +10,7 @@ import baseline.BaselineNeuralNetwork;
 import dataprocessing.CrossValidation;
 import dataprocessing.Protein;
 import dataprocessing.ProteinDataSet;
+import ensemble.NeuralNetEnsemble;
 
 public class ProteinSSClassification {
 	
@@ -19,10 +18,14 @@ public class ProteinSSClassification {
 	BufferedWriter outputFile = null;
 	
 	public ProteinSSClassification(String trainFilename, String testFilename,  int windowSize, int numHiddenLayers,
-			int hiddenLayerSize, int iterations, String outputDir, boolean baseline, boolean traditionalOutput, double decayBy){
+			int hiddenLayerSize, int iterations, String outputDir, boolean baseline, boolean traditionalOutput, double decayBy,
+			boolean ensemble){
 
-
-		if(baseline) {
+		if(ensemble){
+			System.out.println("Creating ensemble");
+			NeuralNetEnsemble NNEnsemble = new NeuralNetEnsemble(trainFilename, testFilename, hiddenLayerSize, windowSize, 5);
+		}
+		else if(baseline) {
 			runBaseline(trainFilename, testFilename, hiddenLayerSize, windowSize, outputDir);
 		}
 		else {
@@ -198,7 +201,7 @@ public class ProteinSSClassification {
 		}
 		
 		ProteinSSClassification classification = new ProteinSSClassification(trainFilename, testFilename, windowSize, 
-					numHiddenLayers, hiddenLayerSize, iterations, outputDir, baseline,traditionalOutput, decayBy);		
+					numHiddenLayers, hiddenLayerSize, iterations, outputDir, baseline,traditionalOutput, decayBy, ensemble);		
 		
 	}
 	

### 2013-12-15 ProteinSSClassification/src/ensemble/NeuralNetEnsemble.java

@@ -30,6 +30,7 @@ public class NeuralNetEnsemble {
 		bagProteins();
 		createNeuralNetworks();
 		
+		classifyTestInstances();
 	}
 	
 	public void classifyTestInstances() {
@@ -84,16 +85,20 @@ public class NeuralNetEnsemble {
 			return STRUCTURE.LOOP;
 		}
 		else if( alpha == loop && alpha == beta) {
+			System.out.println("Equal: all same");
 			// All three are the same, loop is most common so return that
 			return STRUCTURE.LOOP;
 		}
 		else if (alpha == beta) {
+			System.out.println("Equal: alpha and beta");
 			return STRUCTURE.ALPHA;
 		}
 		else if(alpha == loop) {
+			System.out.println("Equal: alpha and loop");
 			return STRUCTURE.LOOP;
 		}
 		else if(beta == loop) {
+			System.out.println("Equal: beta and loop");
 			return STRUCTURE.LOOP;
 		}
 		else {

### 2013-12-15 experiments/runAutoencoderLayerTest_Adam.sh
commiting data from yesterday's runs

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+# Usage: <protein train> <protein test> 
+#                 <window size> <number of hidden layers> <hidden layer size> <iterations> <output directory> <baseline>
+
+trainFilename="/home/vail/Documents/school/cs760/ProteinSSClassification/data/data.train"
+outputDirectory="/home/vail/Documents/school/cs760/ProteinSSClassification/results/deep/layers_adam"
+
+numHiddenLayers=2
+hiddenLayerSize=26 # default value provided by baseline tests
+iterations=5 # increase the number of iterations as we go
+windowSize=13
+baseline="n"
+connectedOutput="n"
+
+
+
+for i in 8 10
+do
+    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
+    echo "Running Deep Network: hidden layer size 26, layer number -> $i" >> "${outputDirectory}/deep.log"
+    java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline $connectedOutput > $outputFile
+    echo "Done" >> "${outputDirectory}/deep.log"
+done

### 2013-12-15 experiments/runBaselineIterationsTest.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# Usage: <protein train> <protein test> 
+#                 <window size> <number of hidden layers> <hidden layer size> <iterations> <output directory> <baseline>
+
+trainFilename="/home/vail/Documents/school/cs760/ProteinSSClassification/data/data.train"
+outputDirectory="/home/vail/Documents/school/cs760/ProteinSSClassification/results/baseline/iterations"
+
+numHiddenLayers=2
+hiddenLayerSize=15 # default value provided by baseline tests
+iterations=5 # increase the number of iterations as we go
+windowSize=13
+baseline="y"
+connectedOutput="n"
+
+# Test the number of hidden layers
+for i in 1 2 5 10
+do
+    outputFile="${outputDirectory}/deep_${numHiddenLayers}nhl_${hiddenLayerSize}hls_${windowSize}ws_${i}iters"
+    echo "Number of iterations -> $i" >> "${outputDirectory}/deep.log"
+    java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $hiddenLayerSize $i $outputDirectory $baseline $connectedOutput > $outputFile
+    echo "Done" >> "${outputDirectory}/deep.log"
+done

### 2013-12-16 experiments/runExperiments.sh
Num layers experiment, 10 iterations.  Results BAD

@@ -12,11 +12,30 @@ iterations=5 # increase the number of iterations as we go
 windowSize=13
 baseline="n"
 
+## Test the number of hidden layers
+#for i in 1 2 3
+#do
+#    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
+#    echo "Running Deep Network: hidden layer size 15, layer number -> $i" >> "${outputDirectory}/deep.log"
+#    java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
+#    echo "Done" >> "${outputDirectory}/deep.log"
+#done
+#
+#hiddenLayerSize=26
+#for i in 1 2 3
+#do
+#    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
+#    echo "Running Deep Network: hidden layer size 26, layer number -> $i" >> "${outputDirectory}/deep.log"
+#    java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
+#    echo "Done" >> "${outputDirectory}/deep.log"
+#done
+
+iterations=10
 # Test the number of hidden layers
 for i in 1 2 3
 do
-    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
-    echo "Running Deep Network: hidden layer size 15, layer number -> $i" >> "${outputDirectory}/deep.log"
+    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws_${iterations}it"
+    echo "Running Deep Network: hidden layer size 15, layer number -> $i, ${iterations} iter" >> "${outputDirectory}/deep.log"
     java -jar deepNetwork.jar $trainFilename "none" $windowSize $i $hiddenLayerSize $iterations $outputDirectory $baseline > $outputFile
     echo "Done" >> "${outputDirectory}/deep.log"
 done
@@ -24,8 +43,8 @@ done
 hiddenLayerSize=26
 for i in 1 2 3
 do
-    outputFile="${outputDirectory}/deep_${i}nhl_${hiddenLayerSize}hls_${windowSize}ws"
[…]

### 2013-12-16 ProteinSSClassification/src/neuralnetwork/test/ThreeLayerNeuralNetworkTest.java
Get tests working again

@@ -18,6 +18,7 @@ public class ThreeLayerNeuralNetworkTest {
 			inputLayer.add(new Unit(false));
 		}
 		NeuralNetworkController netAE = new NeuralNetworkController(inputLayer, 5, null);
+		netAE.setMaxIterations(10000);
 
 		// Input: exactly 1 unit turned on
 		// Teach: alpha if units 0/1 turned on, beta if units 2/3 turned on,
@@ -78,6 +79,7 @@ public class ThreeLayerNeuralNetworkTest {
 		 
 		// Train the system
 		NeuralNetworkController net = new NeuralNetworkController(outAE.get(1), null);
+		net.setMaxIterations(10000);
 		ArrayList<ArrayList<Double>> fedData = new ArrayList<ArrayList<Double>>();
 		for(ArrayList<Double> d : data) {
 			fedData.add(netAE.testHiddenLayer(d));

### 2013-12-16 ProteinSSClassification/src/baseline/BaselineNeuralNetwork.java
tracing through runs

@@ -379,7 +379,7 @@ public class BaselineNeuralNetwork {
 		
 		for(int i = 0; i < allUnits.get(allUnits.size() - 1).size(); i++){
 			//writeOutput(allUnits.get(allUnits.size() - 1).get(i).getValue() + " ");
-			System.out.print(allUnits.get(allUnits.size() - 1).get(i).getValue() + " ");
+			//System.out.print(allUnits.get(allUnits.size() - 1).get(i).getValue() + " ");
 			
 			Unit output = allUnits.get(allUnits.size() - 1).get(i);
 			if(output.getValue() > highestOutput){
@@ -388,7 +388,7 @@ public class BaselineNeuralNetwork {
 			}
 		}
 		//writeOutput("\n");
-		System.out.println();
+		//System.out.println();
 		
 		switch(winner){
 		case 0:

### 2013-12-16 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java

@@ -103,16 +103,32 @@ public class NeuralNetworkController {
 		
 		int iter = 0;
 		int squaredNumWeights = weights.size() * weights.size();
+		int instancePosition = 0;
 		while (squaredSum(weights, weights_old) > TOLERANCE * squaredNumWeights && iter < MAX_ITER) {
-			for(ArrayList<Double> inst : inputValues){
+			/*for(ArrayList<Double> inst : inputValues){
 				setInputs(inst);
 				//printInitialWeights();
 				feedForward();
 				backPropagateAE(inst);
 				//printGraph();
-			}	
+			}
+			*/
+			if(instancePosition == inputValues.size()){
+				instancePosition = 0;
+			}
+			setInputs(inputValues.get(instancePosition));
+			feedForward();
+			backPropagateAE(inputValues.get(instancePosition));
+			instancePosition++;
+			
 			weights_old = weights;
 			weights = getWeights(allUnits.size() - 2);
+			
+			//for(int i = 0; i < weights.size(); i++) {
+			//	System.out.print(weights.get(i) + " ");
+			//}
+			//System.out.println();
+			
 			iter++;
 		}
 		System.out.println("iters to train: " + iter + 

### 2013-12-16 experiments/runBaseline_hiddenLayerSize.sh

@@ -15,7 +15,7 @@ hiddenLayerSize=50
 iterations=5 
 windowSize=13
 baseline="y"
-
+connectedOutput="n"
 
 # Baseline experiments
 count=1
@@ -24,5 +24,5 @@ do
     echo "Running Baseline: hidden layer size -> $i" >> "${outputDirectory}/baseline.log"
     outputFile="${outputDirectory}/baseline_${numHiddenLayers}nhl_${i}hls_${windowSize}ws"
     count=$(($count + 1))
-    java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $i $iterations $outputDirectory $baseline > $outputFile
+    java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $i $iterations $outputDirectory $baseline $connectedOutput > $outputFile
 done

### 2013-12-17 docs/paper.tex
Paper, v1

\documentclass[]{article}
%\usepackage{verbatim}
%\usepackage[dvips]{graphicx}
%\usepackage{pslatex}
\title{Classifying Protein Secondary Structures through Deep Network Learning}
\author{Stephanie deWet and Adam Vail}

\begin{document}
\maketitle

\section{Introduction}
\label{subsec:intro}
Predicting the underlying 3D structure of a protein from its amino acid sequence is a significant problem in structural molecular biology.
An important subproblem of this is to predict complex local structures, or secondary structures, from the amino acid sequence.  
The secondary structure problem has been attacked in a variety of ways, including with x-ray crystallography, biophysical models, and machine learning models.

Within in the field of machine learning, a variety of methods have been applied  over the past twenty years, with varying results. 
The majority of early work was done using standard neural networks with zero or one hidden layers resulting in the vast majority falling below 70\% accuracy.
The recent rise in popularity in deep networks (i.e. neural networks with multiple hidden layers) led us to investigate their accuracy in this problem space.

The common method of encoding amino acids uses large input vectors with an extremely sparse 1-ok-k representation.
Deep networks have the ability to reduce the dimensionality of large sparse feature spaces and are potentially well suited to simplifying the secondary structure problem.

\section{Related Works}
\label{subsec:relatedworks}
One of the earliest studies in this field was conducted by Qian and Sejnowski \cite{Qian} in 1988.
They experimented with using perceptrons and standard back-propagated single hidden layer neural networks.
They achieved results of 63.4\% accuracy which provides us with a reference point for our baselines.

Stolorz et al. combined perceptrons with bayesian methods to probabilistically inject domain knowledge into their model.
The combination of these two methods yielded an improvement over Qian with an accuracy of 64.4\%

Rost and Sander experimented with more complex network structures similar to deep networks.
They created an ensemble of complex networks consisting of two layers of standard single hidden layer neural networks, with the output of the first being the input of the second.
Within the ensemble, each network was trained using varying parameters in order to create separation between the models.
A majority vote was used to produce the final output which was modified in special cases using specific domain knowledge.
This method produced a significant increase in accuracy achieving 70.8\%.

\section{Method}
As alluded to in section \ref{subsec:intro}, our hypothesis is that a deep network has the ability to outperform a standard neural network on the protein secondary structure classification
problem.
Our objective is to find the optimal tuning of the deep network to yield the best predictive accuracy.

\subsection{Deep Networks Consisting of Stacked Autoencoders}
We choose a standard approach to creating a deep network by stacking varying numbers of autoencoders \cite{Hinton} which feed into a set of output units.
An autoencoder is a neural network with a single hidden layer that is trained using backpropagation to reproduce its inputs as outputs.
When there are less hidden units then inputs, the autoencoders enforces a more compact representation of the feature vector.
The process of creating a deep network consists of training a set stacked autoencoders where each successive autoencoder's inputs are the previous autoencoder's hidden layer.

\subsubsection{Input Encoding}
Our training data consists of a set of proteins, each of which is a variable length sequence of amino acids.
Each amino acid in the chain can take on one of twenty discrete values.
Therefore, there are two issues that must be overcome to use an amino acid sequence as input to a neural network: handling variable length inputs and modeling discrete values.

We model the discrete values by using a 1-of-k encoding.
Each amino acid is represented by twenty input nodes, one for each of its potential values.
Of the twenty input nodes, we set the input node corresponding to the amino acid's value to one and set all others to zero.
This was chosen because we did not want to impose an ordering on the values of the amino acid or to make certain values look artificially closer to each other.
Rost and Qian chose a similar encoding.

In order to handle variable length inputs and the fact that protein sequences tend to be quite long, we chose to use a sliding window over the sequence.
For every sliding window the goal is to classify the secondary structure of the amino acid at the center of the window.
The effect of changing the window size changes the impact of neighbors on the central amino acid's classification.
The challenge of using a sliding window is handling the amino acids that reside within half the window size of either end of the sequence.
We have chosen to model the ends of the sequence by setting each nonexistent (due to being off the end of the sequence) amino acid's twenty input nodes all to zero.
This method of using a sliding window was also used by Rost and Qian.

\subsubsection{Output Layer}
Our three classifications are the coarse secondary structures: alpha helix, beta sheet, or loop (known in some of the literature as coil).
We represent each of these possibilities with an output unit, where the highest value of the three output units is chosen in a winner-take-all fashion.
These outputs are attached to the hidden layer of the final autoencoder, and the resultant perceptrons are trained with back propagation.
We use a sigmoid function in all of the hidden and output units.

\subsection{Modifications to the Basic Algorithm}
We have run some tests with each of the following modifications to the deep network.

\subsubsection{Decaying hidden layer size}
Because then input is a large sparse array, a more compact representation is desirable.
We use a decay factor to decrease the size of each stacked hidden layer, incrementally reducing the size of the feature space.
This is designed to find underlying structures in the data.

\subsubsection{Additional Connections to Output Units}
Due to the fact that the autoencoders find a compact representation of the data, we wanted to see whether we were losing valuable information in hidden layers.
We modified the output layer to connect to every input and hidden unit, so that the final layer has inputs from every layer of the deep network.
We then trained the final layer as a giant perceptron, and continued to use a winner-take-all final prediction.

\subsection{Baseline}

\begin{thebibliography}{9}

\bibitem{Hinton}
G.E. Hinton and R. R. Salakhutdinov,
   `` Reducing the Dimensionality of Data with Neural Networks"
   in \emph{Science}, 2006.

\bibitem{Qian}
Ning Qian and Terrence J. Sejnowski,
  `` Predicting the Secondary Structure of Globular Proteins Using Neural Network Models"
  in \emph{J. Mol. Bio.}, 1988.

\bibitem{Rost}
Burkhard Rost and Christ Sander,
  `` Prediction of Protein Secondary Structures at Bettter than 70\% Accuracy"
  in \emph{J. Mol. Bio.}, 1993.

\bibitem{Stolorz}
Paul Stolorz, Alan Lapedes, and Yuan Xia,
   `` Predicting Protein Secondary Structure Using Neural Net and Statistical Methods"
   in \emph{J. Mol. Bio.}, 1992.

\end{thebibliography}

\end{document}

### 2013-12-17 docs/paper.tex
first go of paper, done for the night

@@
 \subsection{Baseline}
+Because the related works in section \ref{subsec:relatedworks} all used domain knowledge in some form, we developed our own baseline model for comparison to our deep network.
+The baseline is a standar neural network using a single hidden layer and trained with back propagation.
+
+\subsubsection{Ensemble of Neural Networks}
+The accuracy gained by Rost from the use of an ensemble inspired us to use an ensemble as well.
+Unfortunately, due to the time complexity of convergence and limitations of physical compute resources, we were only able to develop this for the simpler baseline neural network.
+A majority vote on the outputs of the individual neural networks was used to determine the final output of the ensemble.
 

### 2013-12-17 ProteinSSClassification/src/neuralnetwork/test/AutoencoderTest.java
Update tests to use the minimum required number of iterations

@@ -41,7 +41,8 @@ public class AutoencoderTest {
 		// Print the input values to the console
 		
 		System.out.println("\nHidden layer 1");
-		controller.learnInitialLayer(5, 10000);
+		int numIter = 1500;
+		controller.learnInitialLayer(5, numIter);
 		//controller.printWeights();
 		for (ArrayList<Double> d: controller.getProcessedData()) {
 			controller.getMostRecentLayer().classifyAutoencoder(d);
@@ -60,7 +61,7 @@ public class AutoencoderTest {
 		ArrayList<ArrayList<Double>> newData = controller.feedDataThroughNetwork();  // Outputs of hidden layer
 
 		System.out.println("\nHidden layer 2");		
-		controller.learnHiddenLayer(5, 10000);
+		controller.learnHiddenLayer(5, numIter);
 		
 		//controller.printWeights();		
 		//controller.getMostRecentLayer().printHiddenWeightsFromInput(0);
@@ -73,7 +74,7 @@ public class AutoencoderTest {
 		}
 		
 		System.out.println("\nOutput layer");
-		controller.learnOutputLayer(10000);
+		controller.learnOutputLayer(numIter);
 		
 		//controller.printWeights();		
 		//controller.getMostRecentLayer().printOutputWeightsFromHiddenLayer();		

### 2013-12-17 ProteinSSClassification/src/neuralnetwork/test/PerceptronTest.java

@@ -23,7 +23,7 @@ public class PerceptronTest {
 			inputLayer.add(new Unit(false));
 		}
 		NeuralNetworkController net = new NeuralNetworkController(inputLayer, null);
-		net.setMaxIterations(1000);
+		net.setMaxIterations(100);
 
 		// Input: exactly 1 unit turned on
 		// Teach: alpha if units 0/1 turned on, beta if units 2/3 turned on, loop

### 2013-12-17 ProteinSSClassification/src/classification/ProteinSSClassification.java
changes done at study day

@@ -43,7 +43,8 @@ public class ProteinSSClassification {
 		String date = c.get(Calendar.YEAR)+"-"+c.get(Calendar.MONTH)+"-"+c.get(Calendar.DAY_OF_MONTH);
 		String time = c.get(Calendar.HOUR_OF_DAY)+":"+c.get(Calendar.MINUTE)+":"+c.get(Calendar.SECOND);
 		String outputFilenameBase = date + "_" + time + "_" + numHiddenLayers + "hl_" + hiddenLayerSize + "hls_" + windowSize + "ws";
-				
+		
+		
 		if(testFilename == null) {
 			data = CrossValidation.processData(crossValidationDegree, trainFilename);			
 		}
@@ -70,7 +71,7 @@ public class ProteinSSClassification {
 		pdata.addProteinToTrain(new Protein("YYF", "hhe"));
 		pdata.addProteinListToTest(pdata.getTrain());
 		data.add(pdata);	
-		*/	
+		*/
 		
 		int iter = 1;
 		// Loop through all our data with the given parameters
@@ -84,10 +85,10 @@ public class ProteinSSClassification {
 			*/
 			
 			AutoencoderController controller = new AutoencoderController(d, windowSize, outputFile);
-			controller.learnInitialLayer(hiddenLayerSize, iterations);
+			controller.learnInitialLayer(hiddenLayerSize, 25);
 			int decayAmt = (int)(hiddenLayerSize * decayBy);
-			for(int i = 0; i < numHiddenLayers; i++) {
-				controller.learnHiddenLayer(hiddenLayerSize - i * decayAmt, iterations);
+			for(int i = 0; i < numHiddenLayers - 1; i++) {
+				controller.learnHiddenLayer(hiddenLayerSize - i * decayAmt, 5);
 			}
 			if(traditionalOutput) {
 				controller.learnOutputLayer(iterations);
@@ -186,7 +187,7 @@ public class ProteinSSClassification {
 		if(args[8].equalsIgnoreCase("no") || args[8].equalsIgnoreCase("n")) {
 			traditionalOutput = false;
 		}
-		double decayBy = 1.0;
+		double decayBy = 0;
 		if (args.length == 9) {
 			decayBy = Double.parseDouble(args[9]);
 		}

### 2013-12-17 ProteinSSClassification/src/dataprocessing/CrossValidation.java

@@ -16,10 +16,10 @@ public class CrossValidation {
 		ArrayList<ProteinDataSet> data = new ArrayList<ProteinDataSet>();
 		ProteinDataSet dataSet = new ProteinDataSet();
 		
-//		ArrayList<Protein> train = new ArrayList<Protein>();
-//		train.addAll(readData(trainName).subList(0, 50));
+		ArrayList<Protein> train = new ArrayList<Protein>();
+		train.addAll(readData(trainName).subList(0, 40));
 		
-		dataSet.addProteinListToTrain(readData(trainName));
+		dataSet.addProteinListToTrain(train);
 		dataSet.addProteinListToTest(readData(testName));
 		
 		data.add(dataSet);

### 2013-12-17 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java

@@ -113,6 +113,10 @@ public class NeuralNetworkController {
 			}	
 			weights_old = weights;
 			weights = getWeights(allUnits.size() - 2);
+			for(int i = 0; i < weights.size(); i++){
+				System.out.print(weights.get(i) + " ");
+			}
+			System.out.println();
 			iter++;
 		}
 		System.out.println("iters to train: " + iter + 

### 2013-12-17 ProteinSSClassification/src/classification/ProteinSSClassification.java
Modify code to be ready for more experiments

@@ -43,7 +43,8 @@ public class ProteinSSClassification {
 		String date = c.get(Calendar.YEAR)+"-"+c.get(Calendar.MONTH)+"-"+c.get(Calendar.DAY_OF_MONTH);
 		String time = c.get(Calendar.HOUR_OF_DAY)+":"+c.get(Calendar.MINUTE)+":"+c.get(Calendar.SECOND);
 		String outputFilenameBase = date + "_" + time + "_" + numHiddenLayers + "hl_" + hiddenLayerSize + "hls_" + windowSize + "ws";
-		/*	
+	
+		
 		if(testFilename == null) {
 			data = CrossValidation.processData(crossValidationDegree, trainFilename);			
 		}
@@ -51,8 +52,9 @@ public class ProteinSSClassification {
 			System.out.println("Using UCI data");
 			data = CrossValidation.processData(trainFilename, testFilename);
 		}
-		*/
 		
+
+		/*
 		data = new ArrayList<ProteinDataSet>();
 		ProteinDataSet pdata = new ProteinDataSet();
 		pdata.addProteinToTrain(new Protein("Y", "h"));
@@ -69,7 +71,8 @@ public class ProteinSSClassification {
 
 		pdata.addProteinToTrain(new Protein("YYF", "hhe"));
 		pdata.addProteinListToTest(pdata.getTrain());
-		data.add(pdata);	
+		data.add(pdata);
+		*/	
 		
 		
 		int iter = 1;
@@ -87,6 +90,8 @@ public class ProteinSSClassification {
 			controller.learnInitialLayer(hiddenLayerSize, iterations);
 			int decayAmt = (int)(hiddenLayerSize * decayBy);
 			for(int i = 0; i < numHiddenLayers - 1; i++) {
+				writeOutput("Begin: learn Hidden layer of size " + (hiddenLayerSize - i * decayAmt));
+				System.out.println("Begin: learn Hidden layer of size " + (hiddenLayerSize - i * decayAmt));
 				controller.learnHiddenLayer(hiddenLayerSize - i * decayAmt, iterations);
 			}
 			if(traditionalOutput) {
@@ -186,7 +191,7 @@ public class ProteinSSClassification {
 		if(args[8].equalsIgnoreCase("no") || args[8].equalsIgnoreCase("n")) {
 			traditionalOutput = false;
 		}
-		double decayBy = 1.0;
+		double decayBy = 0.0;
 		if (args.length == 9) {
 			decayBy = Double.parseDouble(args[9]);
 		}

### 2013-12-17 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java

@@ -103,16 +103,16 @@ public class NeuralNetworkController {
 		
 		int iter = 0;
 		int squaredNumWeights = weights.size() * weights.size();
-		int instancePosition = 0;
+		//int instancePosition = 0;
 		while (squaredSum(weights, weights_old) > TOLERANCE * squaredNumWeights && iter < MAX_ITER) {
-			/*for(ArrayList<Double> inst : inputValues){
+			for(ArrayList<Double> inst : inputValues){
 				setInputs(inst);
 				//printInitialWeights();
 				feedForward();
 				backPropagateAE(inst);
 				//printGraph();
 			}
-			*/
+			/*
 			if(instancePosition == inputValues.size()){
 				instancePosition = 0;
 			}
@@ -120,6 +120,7 @@ public class NeuralNetworkController {
 			feedForward();
 			backPropagateAE(inputValues.get(instancePosition));
 			instancePosition++;
+			*/
 			
 			weights_old = weights;
 			weights = getWeights(allUnits.size() - 2);
@@ -127,8 +128,7 @@ public class NeuralNetworkController {
 			//for(int i = 0; i < weights.size(); i++) {
 			//	System.out.print(weights.get(i) + " ");
 			//}
-			//System.out.println();
-			
+			//System.out.println();			
 			iter++;
 		}
 		System.out.println("iters to train: " + iter + 

### 2013-12-17 ProteinSSClassification/src/neuralnetwork/NeuralNetworkController.java
reset autoencoderLearn back to the original way where we check weights after an entire iteration over the inputs

@@ -103,23 +103,14 @@ public class NeuralNetworkController {
 		
 		int iter = 0;
 		int squaredNumWeights = weights.size() * weights.size();
-		int instancePosition = 0;
 		while (squaredSum(weights, weights_old) > TOLERANCE * squaredNumWeights && iter < MAX_ITER) {
-			/*for(ArrayList<Double> inst : inputValues){
+			for(ArrayList<Double> inst : inputValues){
 				setInputs(inst);
 				//printInitialWeights();
 				feedForward();
 				backPropagateAE(inst);
 				//printGraph();
 			}
-			*/
-			if(instancePosition == inputValues.size()){
-				instancePosition = 0;
-			}
-			setInputs(inputValues.get(instancePosition));
-			feedForward();
-			backPropagateAE(inputValues.get(instancePosition));
-			instancePosition++;
 			
 			weights_old = weights;
 			weights = getWeights(allUnits.size() - 2);

### 2013-12-17 ProteinSSClassification/src/dataprocessing/CrossValidation.java
using all data again. was using less before because of memory

@@ -16,10 +16,7 @@ public class CrossValidation {
 		ArrayList<ProteinDataSet> data = new ArrayList<ProteinDataSet>();
 		ProteinDataSet dataSet = new ProteinDataSet();
 		
-		ArrayList<Protein> train = new ArrayList<Protein>();
-		train.addAll(readData(trainName).subList(0, 40));
-		
-		dataSet.addProteinListToTrain(train);
+		dataSet.addProteinListToTrain(readData(trainName));
 		dataSet.addProteinListToTest(readData(testName));
 		
 		data.add(dataSet);

### 2013-12-17 ProteinSSClassification/src/classification/ProteinSSClassification.java
reset initial layer iterations to whatever is passed in on command line

@@ -85,7 +85,7 @@ public class ProteinSSClassification {
 			*/
 			
 			AutoencoderController controller = new AutoencoderController(d, windowSize, outputFile);
-			controller.learnInitialLayer(hiddenLayerSize, 25);
+			controller.learnInitialLayer(hiddenLayerSize, iterations);
 			int decayAmt = (int)(hiddenLayerSize * decayBy);
 			for(int i = 0; i < numHiddenLayers - 1; i++) {
 				controller.learnHiddenLayer(hiddenLayerSize - i * decayAmt, iterations);

### 2013-12-17 ProteinSSClassification/src/classification/ProteinSSClassification.java
Fix decayBy - it was broken

@@ -191,10 +191,8 @@ public class ProteinSSClassification {
 		if(args[8].equalsIgnoreCase("no") || args[8].equalsIgnoreCase("n")) {
 			traditionalOutput = false;
 		}
-		double decayBy = 0.0;
-		if (args.length == 9) {
-			decayBy = Double.parseDouble(args[9]);
-		}
+		double decayBy = Double.parseDouble(args[9]);
+
 		
 		boolean ensemble = false;
 		if(args[10].equalsIgnoreCase("yes") || args[10].equalsIgnoreCase("y")) {

### 2013-12-17 docs/plot.py
added plotting template

@@ -1,0 +1,41 @@
+#!/usr/bin/env python
+import sys
+from matplotlib import pyplot
+
+if __name__ == '__main__':
+
+    input_file = sys.argv[1]
+    output_file = sys.argv[2]
+
+    m = []
+    accuracies = []
+
+    values = open(input_file, "r")
+    for line in values:
+        elements = line.split()
+        m.append(int(elements[0]))
+        accuracies.append(float(elements[1]))
+
+#    n_neg_y = []
+#    n_pos_y = []
+   
+    
+    pyplot.figure()
+#    pyplot.errorbar(m_s, n_accuracies, yerr=[n_neg_y, n_pos_y], linestyle='-', marker='o', label="Naive")
+#    pyplot.errorbar(m_s, t_accuracies, yerr=[t_neg_y, t_pos_y], linestyle='-', marker='o',  label="TAN")
+    pyplot.plot(m, accuracies, linestyle='-', marker='o')
+#    pyplot.plot(m_s, t_accuracies, linestyle='-', marker='o', label="TAN")
+    
+#    for i in range(len(m_s)):
+#        pyplot.text(m_s[i] - 3, n_accuracies[i] + 2, n_accuracies[i])
+#        pyplot.text(m_s[i] - 3, t_accuracies[i] + 2, t_accuracies[i])
+
+    pyplot.legend(loc="upper left")
+
+    pyplot.xlabel("Training Set Size")
+    pyplot.xlim(0, 125)
+    pyplot.ylim(50, 100)
+    pyplot.ylabel("Accuracy (%)")
+    pyplot.title("Naive Bayes Training Set Size vs Accuracy")
+    pyplot.savefig("%s.png" % output_file, bbox_inches='tight')
+    pyplot.show()

### 2013-12-17 experiments/runDecayExperiment.sh
Tests with lots of iterations (long run)

@@ -13,20 +13,20 @@ windowSize=13
 baseline="n"
 
 ## Test the number of hidden layers
-#for i in 0.33 0.25
-#do
-#    outputFile="${outputDirectory}/deep_${numHiddenLayers}nhl_${hiddenLayerSize}hls_${windowSize}ws_${i}decay"
-#    echo "Running Deep Network: hidden layer size 45, decay value $i" >> "${outputDirectory}/deep.log"
-#    java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $hiddenLayerSize $iterations $outputDirectory $baseline $i > $outputFile
-#    echo "Done" >> "${outputDirectory}/deep.log"
-#done
-
-iterations=10
-# Test the number of hidden layers
 for i in 0.33 0.25
 do
     outputFile="${outputDirectory}/deep_${numHiddenLayers}nhl_${hiddenLayerSize}hls_${windowSize}ws_${i}decay"
-    echo "Running Deep Network: hidden layer size 45, decay value $i, num iters $iterations" >> "${outputDirectory}/deep.log"
+    echo "Running Deep Network: hidden layer size 45, decay value $i" >> "${outputDirectory}/deep.log"
     java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $hiddenLayerSize $iterations $outputDirectory $baseline $i > $outputFile
     echo "Done" >> "${outputDirectory}/deep.log"
 done
+
+#iterations=10
+## Test the number of hidden layers
+#for i in 0.33 0.25
+#do
+#    outputFile="${outputDirectory}/deep_${numHiddenLayers}nhl_${hiddenLayerSize}hls_${windowSize}ws_${i}decay"
+#    echo "Running Deep Network: hidden layer size 45, decay value $i, num iters $iterations" >> "${outputDirectory}/deep.log"
+#    java -jar deepNetwork.jar $trainFilename "none" $windowSize $numHiddenLayers $hiddenLayerSize $iterations $outputDirectory $baseline $i > $outputFile
+#    echo "Done" >> "${outputDirectory}/deep.log"
+#done

### 2013-12-17 docs/paper.tex
added verification section

@@
 When there are less hidden units then inputs, the autoencoders enforces a more compact representation of the feature vector.
-The process of creating a deep network consists of training a set stacked autoencoders where each successive autoencoder's inputs are the previous autoencoder's hidden layer.
+The process of creating a deep network consists of training a set stacked autoencoders using stochastic gradient descent where each successive autoencoder's inputs are the previous autoencoder's hidden layer.
 
@@
 We represent each of these possibilities with an output unit, where the highest value of the three output units is chosen in a winner-take-all fashion.
-These outputs are attached to the hidden layer of the final autoencoder, and the resultant perceptrons are trained with back propagation.
+These outputs are attached to the hidden layer of the final autoencoder, and the resultant perceptrons are trained with back propagation using stochastic gradient descent.
 We use a sigmoid function in all of the hidden and output units.
@@
 \subsection{Baseline}
-Because the related works in section \ref{subsec:relatedworks} all used domain knowledge in some form, we developed our own baseline model for comparison to our deep network.
-The baseline is a standar neural network using a single hidden layer and trained with back propagation.
+All previous work in section \ref{subsec:relatedworks} used domain knowledge in some form, we developed our own baseline model for comparison to our deep network.
+The baseline is a standard neural network using a single hidden layer and trained with back propagation using stochastic gradient descent.
 
@@
 A majority vote on the outputs of the individual neural networks was used to determine the final output of the ensemble.
+
+\subsection{Verification}
+All code for the baseline neural network and the deep network were verified by hand using small examples (no more than sequences of length greater than 5).
+Specifically, we traced examples through our perceptron, standard single hidden layer neural network, autoencoder, and stacked autoencoders.
+One detail of note, the small examples that were used for verification consistently required at least 1500 iterations over all the training data in order for the weights to converge.
+Therefore, after many hours of debugging, we are confident the functionality of our learners are correct.
 

### 2013-12-17 postprocessors/clean_for_roc.py
ROC with multiple iterations

@@ -4,93 +4,105 @@ from matplotlib import *
 from matplotlib.pyplot import *
 
 output_folder="docs/results/"  # From base level of repo
+input_prefix="results/deep/long_run/deep_30nhl_30hls_13ws_0.16666d_"
+input_files=[input_prefix + '25it', input_prefix + '50it', input_prefix + '100it']
+plottable_curves_x = []
+plottable_curves_y = []
+desired_class = sys.argv[1]
+#output_file = open(output_folder+sys.argv[3]+"_"+desired_class, "w")
 
-result_file = open(sys.argv[1], "r")
-desired_class = sys.argv[2]
-output_file = open(output_folder+sys.argv[3]+"_"+desired_class, "w")
-output_signs = [-1.0, -1.0, -1.0]
-if desired_class == "LOOP":
-    output_signs[2] = 1.0
-elif desired_class == "BETA":
-    output_signs[1] = 1.0
-else:
-    output_signs[0] = 1.0
-plot_smoothed = False
-if (sys.argv[4]=='y' or sys.argv[4]=='Y' or sys.argv[4]=='yes' or sys.argv[4]=='YES'):
-    plot_smoothed = True
+for i in [0, 1, 2]:
+	result_file = open(input_files[i], "r")
+	output_signs = [-1.0, -1.0, -1.0]
+	if desired_class == "LOOP":
+		output_signs[2] = 1.0
+	elif desired_class == "BETA":
+		output_signs[1] = 1.0
+	else:
+		output_signs[0] = 1.0
+	plot_smoothed = False
+	if (sys.argv[2]=='y' or sys.argv[2]=='Y' or sys.argv[2]=='yes' or sys.argv[2]=='YES'):
+		plot_smoothed = True
 
 
 
-# Raw contains: [alpha_val beta_val loop_val predicted_class actual_class] 
-raw = []
+	# Raw contains: [alpha_val beta_val loop_val predicted_class actual_class] 
+	raw = []
 
-# Parse raw input into raw
-for line in result_file:
-    if "iters" in line:
-        elements = line.split()
-    elif "Too" in line:
-        elements = line.split()
-    elif "Accuracy" in line: 
-        elements = line.split()
-    elif "Percentage" in line:
-        elements = line.split()
-    elif "Prediction" in line:
-        elements = line.split()
-        item.append(elements[1])
-        item.append(elements[3])
-        raw.append(item)
-    else:
-        elements = line.split()
-        item = elements
-   
-#for item in raw:
[…]

### 2013-12-17 postprocessors/clean_for_roc.py
Revert "ROC with multiple iterations"
These graphs are not too informative

This reverts commit c46521127ed90be91145bea44506b97cd8dd9542.

@@ -4,105 +4,93 @@ from matplotlib import *
 from matplotlib.pyplot import *
 
 output_folder="docs/results/"  # From base level of repo
-input_prefix="results/deep/long_run/deep_30nhl_30hls_13ws_0.16666d_"
-input_files=[input_prefix + '25it', input_prefix + '50it', input_prefix + '100it']
-plottable_curves_x = []
-plottable_curves_y = []
-desired_class = sys.argv[1]
-#output_file = open(output_folder+sys.argv[3]+"_"+desired_class, "w")
 
-for i in [0, 1, 2]:
-	result_file = open(input_files[i], "r")
-	output_signs = [-1.0, -1.0, -1.0]
-	if desired_class == "LOOP":
-		output_signs[2] = 1.0
-	elif desired_class == "BETA":
-		output_signs[1] = 1.0
-	else:
-		output_signs[0] = 1.0
-	plot_smoothed = False
-	if (sys.argv[2]=='y' or sys.argv[2]=='Y' or sys.argv[2]=='yes' or sys.argv[2]=='YES'):
-		plot_smoothed = True
+result_file = open(sys.argv[1], "r")
+desired_class = sys.argv[2]
+output_file = open(output_folder+sys.argv[3]+"_"+desired_class, "w")
+output_signs = [-1.0, -1.0, -1.0]
+if desired_class == "LOOP":
+    output_signs[2] = 1.0
+elif desired_class == "BETA":
+    output_signs[1] = 1.0
+else:
+    output_signs[0] = 1.0
+plot_smoothed = False
+if (sys.argv[4]=='y' or sys.argv[4]=='Y' or sys.argv[4]=='yes' or sys.argv[4]=='YES'):
+    plot_smoothed = True
 
 
 
-	# Raw contains: [alpha_val beta_val loop_val predicted_class actual_class] 
-	raw = []
+# Raw contains: [alpha_val beta_val loop_val predicted_class actual_class] 
+raw = []
 
-	# Parse raw input into raw
-	for line in result_file:
-		if "Using" in line:
-			elements = line.split()
-		if "iters" in line:
-			elements = line.split()
-		elif "Too" in line:
-			elements = line.split()
-		elif "Accuracy" in line: 
-			elements = line.split()
-		elif "Percentage" in line:
-			elements = line.split()
-		elif "Prediction" in line:
-			elements = line.split()
-			item.append(elements[1])
-			item.append(elements[3])
-			raw.append(item)
[…]

### 2013-12-17 docs/paper.tex
rounded out the bibliograph

@@
-\documentclass[]{article}
+\documentclass[letterpaper,twocolumn,12pt]{article}
 %\usepackage{verbatim}
-%\usepackage[dvips]{graphicx}
-%\usepackage{pslatex}
+\usepackage[]{graphicx}
 \title{Classifying Protein Secondary Structures through Deep Network Learning}
@@
 \subsection{Modifications to the Basic Algorithm}
+\label{subsec:mods}
 We have run some tests with each of the following modifications to the deep network.
@@
 All code for the baseline neural network and the deep network were verified by hand using small examples (no more than sequences of length greater than 5).
-Specifically, we traced examples through our perceptron, standard single hidden layer neural network, autoencoder, and stacked autoencoders.
+Specifically, we traced examples through a perceptron, standard single hidden layer neural network, autoencoder, and stacked autoencoders verifying their correctness.
 One detail of note, the small examples that were used for verification consistently required at least 1500 iterations over all the training data in order for the weights to converge.
 Therefore, after many hours of debugging, we are confident the functionality of our learners are correct.
+
+\section{Empirical Evaluation}
+As discussed in section \ref{subsec:intro}, the parameters we can optimize are sliding window size, hidden layer size, and number of hidden layers.
+We constructed experiments to evaluate various settings of these parameters using both the deep network as well as the baseline neural network.
+Due to the fact that the baseline consists of a single hidden layer, it converges more quickly that the deep network.
+Therefore, we used results from the baseline to inform our choices of initial parameter settings when testing the deep network of autoencoders.
+The basic algorithm modifications described in \ref{subsec:mods} were then evaluated using the best parameter settings from the other experiments.
+
+\subsection{Dataset}
+The datasets used by Rost and Qian were curated with significant domain knowledge therefore we chose to use a combination of these datasets.
+Qian's dataset is currently hosted by the University of California - Irvine's Machine Learning Repository \cite{uci}.
+We used this data without any modification as it was already divided into a 91 protein training set and a test set of size 15 (resulting in 3500 windows of size 13).
+The Rost data consists of 124 proteins.
+Unfortunately, the exact data is not hosted for public use.
+Therefore, we gathered the listed proteins using the DSSP database \cite{DSSP} to provide labels.
+We followed the procedure in \cite{Rost} to group the eight secondary structure classes provided by DSSP into the three standard classes mentioned above.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{.png}
+\caption{Informative caption here}
+\label{Label here}
+\end{figure}
+
+\section{Discussion}
+
+\section{Future Work}
+
+\section{Conclusion}
 
@@
 G.E. Hinton and R. R. Salakhutdinov,
-   `` Reducing the Dimensionality of Data with Neural Networks"
+   `` Reducing the dimensionality of data with neural networks"
    in \emph{Science}, 2006.
+
+\bibitem{DSSP}
+Joosten RP, Te Beek TAH, Krieger E, Hekkelman ML, Hooft RWW, Schneider R, Sander C, Vriend G,
+   `` A series of PDB related databases for everyday needs"
+   in \emph{NAR}, 2010.
+
+\bibitem{DSSP2}
+Kabsch W, Sander C,
+   `` Dictionary of protein secondary structure: pattern recognition of hydrogen-bonded and geometrical features"
+   in \emph{Biopolymers}, 1983.
 
@@
 Ning Qian and Terrence J. Sejnowski,
-  `` Predicting the Secondary Structure of Globular Proteins Using Neural Network Models"
+  `` Predicting the secondary structure of globular proteins using neural network models"
   in \emph{J. Mol. Bio.}, 1988.
@@
 Burkhard Rost and Christ Sander,
-  `` Prediction of Protein Secondary Structures at Bettter than 70\% Accuracy"
+  `` Prediction of protein secondary structures at bettter than 70\% accuracy"
   in \emph{J. Mol. Bio.}, 1993.
@@
 Paul Stolorz, Alan Lapedes, and Yuan Xia,
-   `` Predicting Protein Secondary Structure Using Neural Net and Statistical Methods"
+   `` Predicting protein secondary structure using neural net and statistical methods"
    in \emph{J. Mol. Bio.}, 1992.
+
+\bibitem{uci}
+   C.L. Blake and C.J. Merz,
+   `` UCI repository of machine learning databases"
+   \textbf{http://archive.ics.uci.edu/ml/datasets/Molecular+Biology+(Protein+Secondary+Structure)}
 

### 2013-12-17 docs/paper.tex
updated paper

@@
 Specifically, we traced examples through a perceptron, standard single hidden layer neural network, autoencoder, and stacked autoencoders verifying their correctness.
-One detail of note, the small examples that were used for verification consistently required at least 1500 iterations over all the training data in order for the weights to converge.
+In these tests, we found that small examples that were used for verification consistently required at least 1500 iterations over all the training data in order for the weights to converge.
+Also, in the single-layer neural network test, we found that if the hidden layer size was too small, the autoencoder may be capable of re-creating the input, and each layer may converge, but the representation would be too compact for the full neural
 Therefore, after many hours of debugging, we are confident the functionality of our learners are correct.
@@
 The datasets used by Rost and Qian were curated with significant domain knowledge therefore we chose to use a combination of these datasets.
-Qian's dataset is currently hosted by the University of California - Irvine's Machine Learning Repository \cite{uci}.
+Qian's dataset is currently hosted by the University of California - Irvine's (UCI) Machine Learning Repository \cite{uci}.
 We used this data without any modification as it was already divided into a 91 protein training set and a test set of size 15 (resulting in 3500 windows of size 13).
@@
 Unfortunately, the exact data is not hosted for public use.
-Therefore, we gathered the listed proteins using the DSSP database \cite{DSSP} to provide labels.
+Therefore, we gathered the listed proteins using the Define Secondary Structure of Proteins (DSSP) database \cite{DSSP} \cite{DSSP2} to provide labels.
 We followed the procedure in \cite{Rost} to group the eight secondary structure classes provided by DSSP into the three standard classes mentioned above.
+We combined the Rost dataset with the Qian training data to create our training dataset, and used the UCI test set as our final test set.
+Our parameter tuning experiments were tested using internal seven- or five- fold cross validation on the training data set.
+(Earlier experiments used seven folds, but we switched to five-fold cross validation later due to time complexity.)
+After we chose our parameter values, our final experiments were trained on the entire training set and then tested on the held-aside test set.
+All plotted points in the following experiments are the average of the accuracies of each fold of the cross validation.
+
+\subsection{Baseline Window Size}
+Window size is important because it determines the number of neighbors that can influence an amino acid's classification.
+Figure \ref{fig:baseline_window_size} shows that a window size of 13 or 17 was better than a smaller window.
+This is to be expected, because a larger window uses results from sets a larger neighborhood around the central amino acid.
+However, increasing the window size sharply increased the number of weights that needed to be trained.
+As a compromise between accuraccy and performance, we chose to use a window size of 13 for all other experiments.
 
@@
 \centering
-\includegraphics[width=65mm]{.png}
-\caption{Informative caption here}
-\label{Label here}
+\includegraphics[width=65mm]{results/baseline/baseline_windowSize.png}
+\caption{The effect of window size on accuracy, using a standard neural network. A window size of 13 was chosen for moving forward.}
+\label{fig:baseline_window_size}
 \end{figure}
 
+\subsection{Baseline Hidden Layer Size}
+Hidden layer size determines how much we compact the new representation of the inputs.
+Figure \ref{fig:baseline_hidden_layer_size} shows that fewer units in the hidden layer represent the data better.
+This shows that because the input data is extremely sparse, an entire window of amino acids can be encoded in very few hidden nodes.
+Due to time complexity and physical computing resources, we couldn't test any networks with more than 200 hidden units, and could only test a handful of networks with more than 100 units.
+Regardless, the data show that smaller numbers of hidden units per layer provide better prediction accuracy.
+Based on these results, in subsequent experiments, we do not use more than 50 hidden units per layer.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/baseline/baseline_hiddenLayerSize.png}
+\caption{The effect of hidden layer size, using a standard neural network.  Hidden layer sizes of less than 50 will be used moving forward.}
+\label{fig:baseline_hidden_layer_size}
+\end{figure}
+
+\subsection{Deep Network}
+In deep network tests, we used results from the baseline tests to set the window size to 13 and to choose a few good hidden layer sizes.
+Figure \ref{fig:deep_num_hidden_layers} justifies our use of multiple hidden layer sizes.
+It shows that a larger hidden layer is necessary when there are many hidden layers in the network.
+This is consistent with the results of our single-layer neural network verification test, where a too-small hidden layer would strongly affect the network performance.  
+In fact, we would expect any issues in early layers to be amplified in later layers, and to be even more obvious in networks with more layers.
+
+The data show that the best accuracy occurs for a six hidden layer network.
+Without domain knowledge, it's not clear what underlying physical structure each hidden layer represents, so we are unsure why some numbers of layers work better than others.
+However, the network performance consistently increases  until it reaches six layers in size, after which point there is a large drop in accuracy.
+We believe this is because six layers are capable of creating the most compact representation of the data the network can achieve, and additional layers simply introduce noise.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/deep/deep_numLayers.png}
+\caption{The effect of hidden layer size and number of hidden layers, using a deep network. The best average accuracy (52.5\%) was achieved using 6 hidden layers with 26 units per layer.}
+\label{fig:deep_num_hidden_layers}
+\end{figure}
+
+\subsection{Decay}
+In all of the deep network tests until this point, the sizes of all the hidden layers in a network were the same.
+We modified our algorithm to use a decay factor, so that each hidden layer was some number of units smaller than the previous layer.
+The first hidden for these tests has nearly fifty hidden units, so that the final layer sizes are not too small.
+
+We found that these tests worked signifcantly worse than the standard deep networks.
+Those tests achieved accuracies of approximately 52.5\%.
+With three hidden layers, an initial hidden layer size of 45 and a decay factor of 25\%, we achieved an average accuracy of \textbf{50.37\%}.
+With three hidden layers, an initial hidden layer size of 45 and a decay factor of 33\%, we achieved an average accuracy of \textbf{50.22\%}.
+
+We hypothesize that this is because the network is being compacted too much at each step, to the point of losing information.
+
 \section{Discussion}
+
+% Insert and discuss ROC curves -- should I put ALPHA, BETA, LOOP on one plot?
+
+As we mentioned in verication, if we used too few iterations for the network to converge, we saw a significant decrease in the performance of the network.
 

### 2013-12-18 docs/paper.tex
Paper - first draft complete aside from conclusions, future work, abstract

@@
 Due to time complexity and physical computing resources, we couldn't test any networks with more than 200 hidden units, and could only test a handful of networks with more than 100 units.
+Also, we did not test any networks with size less than the window size.
 Regardless, the data show that smaller numbers of hidden units per layer provide better prediction accuracy.
@@
 
+\subsection{Best Baseline}
+Our best baseline network used a window size of 13 and a single 15-unit hidden layer to produce an average accuracy of 59.51\%.
+Qian's parameter tuning experiments also showed that different tunings had a relatively small impact on accuracy (less than a percentage point).
+Our parameter tuning experiments (for both baseline and deep network), had a similar outcome.
+
 \subsection{Deep Network}
@@
 Figure \ref{fig:deep_num_hidden_layers} justifies our use of multiple hidden layer sizes.
-It shows that a larger hidden layer is necessary when there are many hidden layers in the network.
+It shows that a larger hidden layer is necessary when there are many hidden layers in the network, or when an autoencoder is used instead of backpropagation.
 This is consistent with the results of our single-layer neural network verification test, where a too-small hidden layer would strongly affect the network performance.
 In fact, we would expect any issues in early layers to be amplified in later layers, and to be even more obvious in networks with more layers.
+Additionally, the network with fifteen hidden units only experiences small improvements additional hidden layers.
+This is expected, because if information is lost in the first layer, additional hidden layers will not be able to recover it, and that information is lost to the network.
 
@@
 \centering
-\includegraphics[width=65mm]{results/deep/deep_numLayers.png}
+\includegraphics[width=65mm]{results/deep/numLayers/deep_numLayers.png}
 \caption{The effect of hidden layer size and number of hidden layers, using a deep network. The best average accuracy (52.5\%) was achieved using 6 hidden layers with 26 units per layer.}
@@
 
+\subsection{Connected Output}
+In order to see if our network was losing information that was available in earlier layers, we tested connecting the output units to every other unit in the graph.
+For a network with 26 hidden units per layer and 3 hidden layers, the result was \textbf{52.20\%} accuracy.
+This is slightly less than the accuracy of our best deep network, with 26 hidden units and 6 layers.
+That implies that the our best deep network did not lose any information between the input and final output layers.
+This modification should not be used with a deeper network, because it makes the final output layer very costly to train because of the explosion in size of the perceptron.
+
 \section{Discussion}
+% Determined optimal params - now use test set too
+After completing the parameter tuning experiments above, we chose optimal parameter settings.
+Networks using these parameters were trained using the entire training set and tested with the held-aside test set.
 
+\subsection{Deep Networks}
+The final deep network has a window size of 13, and 6 hidden layers of size 26.
+Its accuracy is ___\%.
+
+% FIXME how ROC curve created
+We created ROC curves to measure the performance of the deep network.
+Confidence was calculated by subtracting the output values for the two losing nodes from the winning output values.
+Because there are multiple classifications, we can't use the basic 2-classification ROC curve format.
+Instead, we chose to use a one-versus-many ROC curve, which shows the false positive and true positive rates for each class.
+
+% discuss concerns 
+% - not as good as the baseline
+% - why?
+% -  not enough iters (verif section) extremely complex
+
+%FIXME FIGURE HERE.
+
+Despite the fact that the accuracy is lower than the baseline, the ROC curves in \ref{fig:roc_alpha} show that for predictions with high confidence, our network predicts the secondary structure well.
+
+\subsection{Ensembles}
+Our best deep network performs signifcantly worse than the baseline neural network, so a natural modification is to try using ensembles of networks.
+Since deep networks require significant time complexity, we were unable to train an ensemble of deep networks.
+However, we were able to train an ensemble of simple neural networks with one hidden layer.
+%Figure \ref{fig:baseline_ensemble} shows that 
+An ensemble of three neural networks has an average accuracy of 61.65\%.
+The best baseline neural network settings, as discussed in section ______,  produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
+So, using an ensemble clearly improves the performance of the baseline network, and we hypothesize that deep networks would also benefit from this technique.
+
+%\begin{figure}[ht!]
+%\centering
+%\includegraphics[width=65mm]{results/ensemble/ensemble.png}
+%\caption{The effect of different numbers of neural networks on ensemble average accuracy across five runs, using standard neural networks.}
+%\label{fig:baseline_ensemble}
+%\end{figure}
+
+%best baseline: 62.72
+%best ensemble: 63
+%best avg: 3 networks 61.65
+
+\subsection{Final Comments on Stacked Autoencoders}
+As we mentioned in verification, if we used too few iterations for the network to converge, we saw a significant decrease in the performance.
+We hypothesize that due to the complexity of the data, our final deep network has not converged and that is why its accuracy is so much worse than the baseline.
+We are confident that with significantly longer training time (on the order of weeks), the deep network would fully converge.
+
+We still believe that deep networks have the ability to solve the secondary structure problem, but due to the time complexity of learning an effective deep network, other learning methods may be more practical.
+
+% describe results
 % Insert and discuss ROC curves -- should I put ALPHA, BETA, LOOP on one plot?
+% combine into 1????
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/ensemble/ROC_ALPHA.png}
+\caption{}
+\label{fig:roc_alpha}
+\end{figure}
 
-As we mentioned in verication, if we used too few iterations for the network to converge, we saw a significant decrease in the performance of the network.
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/ensemble/ROC_BETA.png}
+\caption{}
+\label{fig:roc_beta}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/ensemble/ROC_LOOP.png}
+\caption{}
+\label{fig:roc_loop}
+\end{figure}
+
+\subsection{Lessons Learned}
+In the process of working on this project, we learned that it is very difficult to understand the re-representations produced by the hidden layers of a large multi-layer network.
+The combination of a complicated network and little domain knowledge made it especially challenging to tell whether these re-representations were useful.
+We also learned that deep networks need to be tested very carefully, since they have so many interconnections and weights.
+Overall, we feel that this was a valuable experience, gaining experience in neural networks.
 
 \section{Future Work}
+%TODO
+Ensembles of autoencoders
+Other things than sigmoid
+- sparse autoencoder
 
 \section{Conclusion}
+%TODO
+
+%TODO Create Abstract
 

### 2013-12-18 docs/paper.tex
Fix latex bugs

@@
 The final deep network has a window size of 13, and 6 hidden layers of size 26.
-Its accuracy is ___\%.
+Its accuracy is %___\%. FIXME
 
-% FIXME how ROC curve created
 We created ROC curves to measure the performance of the deep network.
@@
 An ensemble of three neural networks has an average accuracy of 61.65\%.
-The best baseline neural network settings, as discussed in section ______,  produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
+The best baseline neural network settings, as discussed in section %______, FIXME
+produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
 So, using an ensemble clearly improves the performance of the baseline network, and we hypothesize that deep networks would also benefit from this technique.
@@
 \centering
-\includegraphics[width=65mm]{results/ensemble/ROC_ALPHA.png}
+\includegraphics[width=65mm]{results/ROC_ALPHA.png}
 \caption{}
@@
 \centering
-\includegraphics[width=65mm]{results/ensemble/ROC_BETA.png}
+\includegraphics[width=65mm]{results/ROC_BETA.png}
 \caption{}
@@
 \centering
-\includegraphics[width=65mm]{results/ensemble/ROC_LOOP.png}
+\includegraphics[width=65mm]{results/ROC_LOOP.png}
 \caption{}

### 2013-12-18 docs/paper.tex
ROC curve all in one
Fix weblink in reference

@@
 \centering
-\includegraphics[width=65mm]{results/ROC_ALPHA.png}
+\includegraphics[width=65mm]{results/ROC_all.png}
 \caption{}
-\label{fig:roc_alpha}
-\end{figure}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=65mm]{results/ROC_BETA.png}
-\caption{}
-\label{fig:roc_beta}
-\end{figure}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=65mm]{results/ROC_LOOP.png}
-\caption{}
-\label{fig:roc_loop}
+\label{fig:roc}
 \end{figure}
@@
    `` UCI repository of machine learning databases"
-   \textbf{http://archive.ics.uci.edu/ml/datasets/Molecular+Biology+(Protein+Secondary+Structure)}
+   \textbf{http://archive.ics.uci.edu/ml/datasets/}
+   \textbf{Molecular+Biology+(Protein+}
+   \textbf{Secondary+Structure)}
 

### 2013-12-18 experiments/runFinalDeepNet.sh
Final deep net data

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+# Usage: <protein train> <protein test> 
+#                 <window size> <number of hidden layers> <hidden layer size> <iterations> <output directory> <baseline>
+
+outputDirectory="/mnt/ws/home/sdewet/760-project/ProteinSSClassification/results/deep"
+trainFilename="/mnt/ws/home/sdewet/760-project/ProteinSSClassification/data/data.train"
+testFilename="/mnt/ws/home/sdewet/760-project/ProteinSSClassification/data/uci/uci_protein_data.train"
+
+numHiddenLayers=6
+hiddenLayerSize=26 # default value provided by baseline tests
+iterations=$1 # increase the number of iterations as we go
+windowSize=13
+baseline="n"
+decay=0
+traditionalOutput="y"
+ensemble="n"
+numEnsemble=0
+
+## Test the number of hidden layers
+outputFile="${outputDirectory}/deep_${numHiddenLayers}nhl_${hiddenLayerSize}hls_${windowSize}ws_${iterations}it"
+echo "Running Deep Network: hidden layer size $numHiddenLayers $hiddenLayerSize" >> "${outputDirectory}/deep.log"
+java -jar deepNetwork_decayFix.jar $trainFilename $testFilename $windowSize $numHiddenLayers $hiddenLayerSize $iterations $outputDirectory $baseline $traditionalOutput $decay $ensemble $numEnsemble > $outputFile
+echo "Done" >> "${outputDirectory}/deep.log"

### 2013-12-18 docs/paper.tex
all data and graphs in paper

@@
 \subsection{Best Baseline}
+\label{subsec:bestbaseline}
 Our best baseline network used a window size of 13 and a single 15-unit hidden layer to produce an average accuracy of 59.51\%.
@@
 The final deep network has a window size of 13, and 6 hidden layers of size 26.
-Its accuracy is %___\%. FIXME
+When training the final network, we observed that an increase in training time significantly improved the results. 
+We trained with 25, 50, and 100 iterations over the entire training set, achieving 52.26\%, 52.4\%, and 53.21\% accuracy respectively.
+The 100 iteration network does almost a full percentage point better than the best parameter tuning experiment, which was trained with only 10 iterations due to time demands of cross validation.
 
-We created ROC curves to measure the performance of the deep network.
+We created ROC curves to measure the performance of the best deep network.
 Confidence was calculated by subtracting the output values for the two losing nodes from the winning output values.
 Because there are multiple classifications, we can't use the basic 2-classification ROC curve format.
-Instead, we chose to use a one-versus-many ROC curve, which shows the false positive and true positive rates for each class.
+Instead, we chose to use a one-versus-all ROC curve, which shows the false positive and true positive rates for each class.
 
-% discuss concerns 
-% - not as good as the baseline
-% - why?
-% -  not enough iters (verif section) extremely complex
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=65mm]{results/ROC_all_final_deep_net_100it.png}
+\caption{}
+\label{fig:roc}
+\end{figure}
 
-%FIXME FIGURE HERE.
-
-Despite the fact that the accuracy is lower than the baseline, the ROC curves in \ref{fig:roc_alpha} show that for predictions with high confidence, our network predicts the secondary structure well.
+Despite the fact that the accuracy is lower than the baseline, all three ROC curves in \ref{fig:roc} show that the deep network's performance is better than simply guessing.
+Also note that for predictions with high confidence, our network predicts the secondary structure well, particularily BETA classifications.
 
@@
 However, we were able to train an ensemble of simple neural networks with one hidden layer.
-%Figure \ref{fig:baseline_ensemble} shows that 
 An ensemble of three neural networks has an average accuracy of 61.65\%.
-The best baseline neural network settings, as discussed in section %______, FIXME
+The best baseline neural network settings, as discussed in section \ref{subsec:bestbaseline},
 produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
@@
 As we mentioned in verification, if we used too few iterations for the network to converge, we saw a significant decrease in the performance.
-We hypothesize that due to the complexity of the data, our final deep network has not converged and that is why its accuracy is so much worse than the baseline.
+Due to the complexity of the data, our final deep network has not fully converged and that is why its accuracy is so much worse than the baseline.
 We are confident that with significantly longer training time (on the order of weeks), the deep network would fully converge.
@@
 We still believe that deep networks have the ability to solve the secondary structure problem, but due to the time complexity of learning an effective deep network, other learning methods may be more practical.
-
-% describe results
-% Insert and discuss ROC curves -- should I put ALPHA, BETA, LOOP on one plot?
-% combine into 1????
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=65mm]{results/ROC_all.png}
-\caption{}
-\label{fig:roc}
-\end{figure}
 

### 2013-12-18 docs/paper.tex
Paper first draft complete

@@
 \maketitle
+
+\begin{abstract}
+Classifying the secondary structures of proteins is a well-known problem in molecular biology.
+Our goal is explore the potential of applying deep learning methods to this domain.
+We performed an empirical evaluation of deep network parameters on a network of stacked autoencoders.
+We found that our best performance of 53.2\% resulted from parameters of window size 13, 6 hidden layers, and 26 units per hidden layer.
+Unfortunately, this underperforms as compared to other learning methods.
+\end{abstract}
 
@@
 produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
-So, using an ensemble clearly improves the performance of the baseline network, and we hypothesize that deep networks would also benefit from this technique.
+So, using an ensemble clearly improves the performance of the baseline network, and we hypothesize that deep networks would also benefit from this technique, but leave it to future work.
 
@@
 \section{Future Work}
-%TODO
-Ensembles of autoencoders
-Other things than sigmoid
-- sparse autoencoder
+We considered additional modifications to the basic algorithm that we were unable to test due to time constraints.
+We would have liked to experiment with additional functions for our hidden and output units, rather than only using a sigmoid.
+One interesting modification would be using a function that further encourages sparsity by using a penalty term \cite{ng}.
 
 \section{Conclusion}
-%TODO
-
-%TODO Create Abstract
+We explored a deep network's ability to classsify secondary structures in proteins.
+We created a baseline of a standard single-hidden layer neural network for comparison.
+We did a parameter tuning study, and found that the best parameters for the deep network were a window size of 13 and 6 hidden layers with 26 units per layer, resulting in a final accuracy of 53.2\%.
+We found that our deep networks were unable to achieve the accuracy of a standard neural network.
 
@@
    in \emph{Biopolymers}, 1983.
+
+\bibitem{Ng}
+Andrew Ng,
+   `` CS294 Lecture notes: Sparse autoencoder"
+   \textbf{www.stanford.edu/class/cs294a/}
+   \textbf{sparseAutoencoder.pdf}
 
@@
    \textbf{http://archive.ics.uci.edu/ml/datasets/}
-   \textbf{Molecular+Biology+(Protein+}
-   \textbf{Secondary+Structure)}
+   \textbf{Molecular+Biology+\}
+   \textbf{(Protein+Secondary+Structure)}
 

### 2013-12-18 docs/paper.tex
Fix bracket
almost done editing

@@
 Classifying the secondary structures of proteins is a well-known problem in molecular biology.
-Our goal is explore the potential of applying deep learning methods to this domain.
+Our goal is to explore the potential of applying deep learning methods to this domain.
 We performed an empirical evaluation of deep network parameters on a network of stacked autoencoders.
@@
 
-Within in the field of machine learning, a variety of methods have been applied over the past twenty years, with varying results. 
-The majority of early work was done using standard neural networks with zero or one hidden layers resulting in the vast majority falling below 70\% accuracy.
+Within the field of machine learning, a variety of methods have been applied over the past twenty years, with varying results. 
+Early work was done using standard neural networks with zero or one hidden layers resulting in the vast majority falling below 70\% accuracy.
 The recent rise in popularity in deep networks (i.e. neural networks with multiple hidden layers) led us to investigate their accuracy in this problem space.
 
-The common method of encoding amino acids uses large input vectors with an extremely sparse 1-ok-k representation.
+The common method of encoding amino acids uses large input vectors with an extremely sparse 1-of-k representation.
 Deep networks have the ability to reduce the dimensionality of large sparse feature spaces and are potentially well suited to simplifying the secondary structure problem.
@@
 One of the earliest studies in this field was conducted by Qian and Sejnowski \cite{Qian} in 1988.
-They experimented with using perceptrons and standard back-propagated single hidden layer neural networks.
-They achieved results of 63.4\% accuracy which provides us with a reference point for our baselines.
+They used perceptrons and standard back-propagated single hidden layer neural networks to achieve results of 63.4\% accuracy, providing us with a reference point for our baselines.
 
-Stolorz et al. combined perceptrons with bayesian methods to probabilistically inject domain knowledge into their model.
-The combination of these two methods yielded an improvement over Qian with an accuracy of 64.4\%
+Stolorz et al. combined perceptrons with Bayesian methods to probabilistically inject domain knowledge into their model \cite{Stolorz}.
+The combination of these two methods yielded an improvement over Qian with an accuracy of 64.4\%.
 
-Rost and Sander experimented with more complex network structures similar to deep networks.
+Rost and Sander experimented with more complex network structures similar to deep networks \cite{Rost}.
 They created an ensemble of complex networks consisting of two layers of standard single hidden layer neural networks, with the output of the first being the input of the second.
@@
 A majority vote was used to produce the final output which was modified in special cases using specific domain knowledge.
-This method produced a significant increase in accuracy achieving 70.8\%.
+This method produced a significant increase in performance, achieving 70.8\% in accuracy.
 
@@
 \subsection{Deep Networks Consisting of Stacked Autoencoders}
-We choose a standard approach to creating a deep network by stacking varying numbers of autoencoders \cite{Hinton} which feed into a set of output units.
+We choose a standard approach to creating a deep network by stacking varying numbers of autoencoders \cite{Hinton}, which feed into a set of output units.
 An autoencoder is a neural network with a single hidden layer that is trained using backpropagation to reproduce its inputs as outputs.
-When there are less hidden units then inputs, the autoencoders enforces a more compact representation of the feature vector.
-The process of creating a deep network consists of training a set stacked autoencoders using stochastic gradient descent where each successive autoencoder's inputs are the previous autoencoder's hidden layer.
+When there are less hidden units then inputs, the autoencoders enforce a more compact representation of the feature vector.
+Our deep network consists of a set of stacked autoencoders trained using stochastic gradient descent where each successive autoencoder's inputs are the previous autoencoder's hidden layer.
 
 \subsubsection{Input Encoding}
-Our training data consists of a set of proteins, each of which is a variable length sequence of amino acids.
+The training data consists of a set of proteins, each of which is a variable length sequence of amino acids.
 Each amino acid in the chain can take on one of twenty discrete values.
@@
 For every sliding window the goal is to classify the secondary structure of the amino acid at the center of the window.
-The effect of changing the window size changes the impact of neighbors on the central amino acid's classification.
+Changing the window size affects the impact of neighbors on the central amino acid's classification.
+This sliding window technique was also used by Rost and Qian.
+
 The challenge of using a sliding window is handling the amino acids that reside within half the window size of either end of the sequence.
 We have chosen to model the ends of the sequence by setting each nonexistent (due to being off the end of the sequence) amino acid's twenty input nodes all to zero.
-This method of using a sliding window was also used by Rost and Qian.
 
 \subsubsection{Output Layer}
-Our three classifications are the coarse secondary structures: alpha helix, beta sheet, or loop (known in some of the literature as coil).
+Our three classifications are the coarse secondary structures: alpha helix, beta sheet, and loop (also known as coil).
 We represent each of these possibilities with an output unit, where the highest value of the three output units is chosen in a winner-take-all fashion.
 These outputs are attached to the hidden layer of the final autoencoder, and the resultant perceptrons are trained with back propagation using stochastic gradient descent.
-We use a sigmoid function in all of the hidden and output units.
+We use a sigmoid activation function for all units throughout the network.
 
@@
 \subsubsection{Decaying hidden layer size}
-Because then input is a large sparse array, a more compact representation is desirable.
-We use a decay factor to decrease the size of each stacked hidden layer, incrementally reducing the size of the feature space.
-This is designed to find underlying structures in the data.
+The input is a large sparse array, therefore a more compact representation is desirable.
+We use a decay factor to decrease the size of each stacked hidden layer, rapidly reducing the size of the feature space.
 
 \subsubsection{Additional Connections to Output Units}
-Due to the fact that the autoencoders find a compact representation of the data, we wanted to see whether we were losing valuable information in hidden layers.
-We modified the output layer to connect to every input and hidden unit, so that the final layer has inputs from every layer of the deep network.
-We then trained the final layer as a giant perceptron, and continued to use a winner-take-all final prediction.
+Due to the fact that the autoencoders find a compact representation of the data, we wanted to evaluate whether we were losing valuable information in the hidden layers.
+The output layer was modified to take all other units in the network as input.
+The final layer was then trained as a giant perceptron and continued to use a winner-take-all final predication.
 
 \subsection{Baseline}
-All previous work in section \ref{subsec:relatedworks} used domain knowledge in some form, we developed our own baseline model for comparison to our deep network.
+All related work in section \ref{subsec:relatedworks} used domain knowledge that is unavailable to us, therefore we developed our own baseline model for comparison to our deep network.
 The baseline is a standard neural network using a single hidden layer and trained with back propagation using stochastic gradient descent.
@@
 The accuracy gained by Rost from the use of an ensemble inspired us to use an ensemble as well.
+We used bagging at a protein granularity to create unique training sets for each separate model.
+A majority vote on the outputs of the individual neural networks was used to determine the final output.
 Unfortunately, due to the time complexity of convergence and limitations of physical compute resources, we were only able to develop this for the simpler baseline neural network.
-A majority vote on the outputs of the individual neural networks was used to determine the final output of the ensemble.
 
 \subsection{Verification}
-All code for the baseline neural network and the deep network were verified by hand using small examples (no more than sequences of length greater than 5).
+\label{subsec:verification}
+All code for the baseline neural network and the deep network was verified by hand using small examples (amino acid sequences of length less than five).
 Specifically, we traced examples through a perceptron, standard single hidden layer neural network, autoencoder, and stacked autoencoders verifying their correctness.
-In these tests, we found that small examples that were used for verification consistently required at least 1500 iterations over all the training data in order for the weights to converge.
-Also, in the single-layer neural network test, we found that if the hidden layer size was too small, the autoencoder may be capable of re-creating the input, and each layer may converge, but the representation would be too compact for the full neural
-Therefore, after many hours of debugging, we are confident the functionality of our learners are correct.
+After many hours of debugging, we are confident the functionality of our learners are correct.
+
+We found that the small examples that were used for verification consistently required at least 1500 iterations over the small set of training data in order for the weights to converge.
+Also, if the hidden layer size was too small, each layer may converge, but the representation would be too compact for the full network to accurately predict the structure.
 
@@
 As discussed in section \ref{subsec:intro}, the parameters we can optimize are sliding window size, hidden layer size, and number of hidden layers.
-We constructed experiments to evaluate various settings of these parameters using both the deep network as well as the baseline neural network.
-Due to the fact that the baseline consists of a single hidden layer, it converges more quickly that the deep network.
+We constructed experiments to evaluate various settings of these parameters using both the deep network and the baseline neural network.
+Due to the fact that the baseline consists of a single hidden layer, it converges more quickly than the deep network.
 Therefore, we used results from the baseline to inform our choices of initial parameter settings when testing the deep network of autoencoders.
@@
 \subsection{Dataset}
-The datasets used by Rost and Qian were curated with significant domain knowledge therefore we chose to use a combination of these datasets.
+We chose to use a combination of the datasets created by Rost and Qian because they were curated with significant domain knowledge.
 Qian's dataset is currently hosted by the University of California - Irvine's (UCI) Machine Learning Repository \cite{uci}.
-We used this data without any modification as it was already divided into a 91 protein training set and a test set of size 15 (resulting in 3500 windows of size 13).
-The Rost data consists of 124 proteins.
+We used this data without any modification as it was already divided into a 91 protein training set and a 15 protein test set.
+While the test set is somewhat small, it results in approximately 3500 windows of size 13.
+The Rost data consists of a total of 124 proteins.
 Unfortunately, the exact data is not hosted for public use.
-Therefore, we gathered the listed proteins using the Define Secondary Structure of Proteins (DSSP) database \cite{DSSP} \cite{DSSP2} to provide labels.
-We followed the procedure in \cite{Rost} to group the eight secondary structure classes provided by DSSP into the three standard classes mentioned above.
-We combined the Rost dataset with the Qian training data to create our training dataset, and used the UCI test set as our final test set.
-Our parameter tuning experiments were tested using internal seven- or five- fold cross validation on the training data set.
-(Earlier experiments used seven folds, but we switched to five-fold cross validation later due to time complexity.)
-After we chose our parameter values, our final experiments were trained on the entire training set and then tested on the held-aside test set.
+Therefore, we gathered the listed proteins and their corresponding labels using the Define Secondary Structure of Proteins (DSSP) database \cite{DSSP} \cite{DSSP2}.
+We followed the procedure in \cite{Rost} to group the eight secondary structure classes provided by DSSP into the three standard classes mentioned previously.
+We combined all proteins from Rost with Qian's training proteins to create our training data set.
+
+Our parameter tuning experiments were tested using internal seven- or five- fold cross validation on the training dataset.
+Earlier experiments used seven folds, but we switched to five-fold cross validation later due to time complexity.
+After we chose our parameter values, our final experiments were trained on the entire training set and then tested on a held-aside dataset consisting of Qian's test proteins.
 All plotted points in the following experiments are the average of the accuracies of each fold of the cross validation.
@@
 \subsection{Baseline Window Size}
-Window size is important because it determines the number of neighbors that can influence an amino acid's classification.
-Figure \ref{fig:baseline_window_size} shows that a window size of 13 or 17 was better than a smaller window.
-This is to be expected, because a larger window uses results from sets a larger neighborhood around the central amino acid.
+Window size is important since it determines the number of neighbors that can influence an amino acid's classification.
+Figure \ref{fig:baseline_window_size} shows that a window size of 13 or 17 is better than a smaller window.
+This is to be expected, because a larger window uses results from sets in a larger neighborhood around the central amino acid.
 However, increasing the window size sharply increased the number of weights that needed to be trained.
@@
 \subsection{Baseline Hidden Layer Size}
-Hidden layer size determines how much we compact the new representation of the inputs.
-Figure \ref{fig:baseline_hidden_layer_size} shows that fewer units in the hidden layer represent the data better.
+Hidden layer size determines the degree of compaction of the new representation of the inputs.
+Figure \ref{fig:baseline_hidden_layer_size} shows that fewer units in the hidden layer provides a better representation of the data.
 This shows that because the input data is extremely sparse, an entire window of amino acids can be encoded in very few hidden nodes.
@@
 Also, we did not test any networks with size less than the window size.
-Regardless, the data show that smaller numbers of hidden units per layer provide better prediction accuracy.
+Regardless, the data show that smaller numbers of hidden units per layer provide better predictive accuracy.
 Based on these results, in subsequent experiments, we do not use more than 50 hidden units per layer.
@@
 \subsection{Deep Network}
-In deep network tests, we used results from the baseline tests to set the window size to 13 and to choose a few good hidden layer sizes.
-Figure \ref{fig:deep_num_hidden_layers} justifies our use of multiple hidden layer sizes.
-It shows that a larger hidden layer is necessary when there are many hidden layers in the network, or when an autoencoder is used instead of backpropagation.
-This is consistent with the results of our single-layer neural network verification test, where a too-small hidden layer would strongly affect the network performance.
-In fact, we would expect any issues in early layers to be amplified in later layers, and to be even more obvious in networks with more layers.
-Additionally, the network with fifteen hidden units only experiences small improvements additional hidden layers.
-This is expected, because if information is lost in the first layer, additional hidden layers will not be able to recover it, and that information is lost to the network.
+In deep network tests we used a window size of 13 and hidden layers sizes of less than 50.
+Figure \ref{fig:deep_num_hidden_layers} shows that a larger hidden layer is necessary when stacked autoencoders are used due to the lack of network wide backpropagation.
+This is consistent with the results of our single-layer neural network verification test, where a too-small hidden layer would strongly effect the network performance.
+In fact, we would expect any issues in early layers to be amplified in deeper networks.
+The networks with fifteen hidden units only experience small improvements with additional hidden layers.
+This is unsurprising, because if information is lost in the first layer, additional hidden layers will not be able to recover it, leading to lost information in the network.
 
-The data show that the best accuracy occurs for a six hidden layer network.
+The data show that the best accuracy of 52.5\% occurs for a six hidden layer network with layer sizes of 26 units.
 Without domain knowledge, it's not clear what underlying physical structure each hidden layer represents, so we are unsure why some numbers of layers work better than others.
@@
 We modified our algorithm to use a decay factor, so that each hidden layer was some number of units smaller than the previous layer.
-The first hidden for these tests has nearly fifty hidden units, so that the final layer sizes are not too small.
+The first hidden layer for these tests has 45 hidden units, so that the final layer sizes are not too small.
 
 We found that these tests worked signifcantly worse than the standard deep networks.
-Those tests achieved accuracies of approximately 52.5\%.
-With three hidden layers, an initial hidden layer size of 45 and a decay factor of 25\%, we achieved an average accuracy of \textbf{50.37\%}.
-With three hidden layers, an initial hidden layer size of 45 and a decay factor of 33\%, we achieved an average accuracy of \textbf{50.22\%}.
+With three hidden layers and a decay factor of .25, we achieved an average accuracy of 50.37\%.
+With three hidden layers and a decay factor of .33, we achieved an average accuracy of 50.22\%.
 
@@
 \subsection{Connected Output}
-In order to see if our network was losing information that was available in earlier layers, we tested connecting the output units to every other unit in the graph.
-For a network with 26 hidden units per layer and 3 hidden layers, the result was \textbf{52.20\%} accuracy.
+In order to evaluate whether our network was losing information that was available in earlier layers, we connected the output units to every other unit in the graph.
+For a network with 26 hidden units per layer and 3 hidden layers, the result was 52.20\% accuracy.
 This is slightly less than the accuracy of our best deep network, with 26 hidden units and 6 layers.
 That implies that the our best deep network did not lose any information between the input and final output layers.
-This modification should not be used with a deeper network, because it makes the final output layer very costly to train because of the explosion in size of the perceptron.
+This modification should not be used with a deeper network because it makes the final output layer very costly to train due to the explosion in size of the perceptron.
 
-\section{Discussion}
+\section{Final Results and Discussion}
 % Determined optimal params - now use test set too
@@
 \subsection{Deep Networks}
-The final deep network has a window size of 13, and 6 hidden layers of size 26.
-When training the final network, we observed that an increase in training time significantly improved the results. 
+The final deep network has a window size of 13 and 6 hidden layers of size 26.
+When training the final network we observed that an increase in training time significantly improved the results. 
 We trained with 25, 50, and 100 iterations over the entire training set, achieving 52.26\%, 52.4\%, and 53.21\% accuracy respectively.
-The 100 iteration network does almost a full percentage point better than the best parameter tuning experiment, which was trained with only 10 iterations due to time demands of cross validation.
+The 100 iteration network does almost a full percentage point better than the best parameter tuning experiment, which was trained with only 10 iterations due to the time demands of cross validation.
 
 We created ROC curves to measure the performance of the best deep network.
-Confidence was calculated by subtracting the output values for the two losing nodes from the winning output values.
-Because there are multiple classifications, we can't use the basic 2-classification ROC curve format.
+Confidence was calculated by subtracting the output values for the two losing nodes from the winning output value.
+Due to multiple classifications, we couldn't use the basic 2-classification ROC curve format.
 Instead, we chose to use a one-versus-all ROC curve, which shows the false positive and true positive rates for each class.
@@
 \includegraphics[width=65mm]{results/ROC_all_final_deep_net_100it.png}
-\caption{}
+\caption{ROC curves for each classification shows that the model performs better than simply guessing}
 \label{fig:roc}
@@
 
-Despite the fact that the accuracy is lower than the baseline, all three ROC curves in \ref{fig:roc} show that the deep network's performance is better than simply guessing.
-Also note that for predictions with high confidence, our network predicts the secondary structure well, particularily BETA classifications.
+Despite the fact that the accuracy is lower than the baseline, all three ROC curves in figure \ref{fig:roc} show that the deep network's performance is better than simply guessing.
+Also note that for predictions with high confidence, our network predicts the secondary structure well, particularily beta classifications.
 
 \subsection{Ensembles}
-Our best deep network performs signifcantly worse than the baseline neural network, so a natural modification is to try using ensembles of networks.
+A natural modification to improve performance is to use ensembles of networks.
 Since deep networks require significant time complexity, we were unable to train an ensemble of deep networks.
-However, we were able to train an ensemble of simple neural networks with one hidden layer.
+However, we were able to train an ensemble of simple neural networks with a single thirty unit hidden layer.
 An ensemble of three neural networks has an average accuracy of 61.65\%.
@@
 produced an average accuracy of 59.51\%, which is two percentage points lower than the ensemble.
-So, using an ensemble clearly improves the performance of the baseline network, and we hypothesize that deep networks would also benefit from this technique, but leave it to future work.
+Using an ensemble clearly improves the performance of the baseline network.
+We hypothesize that deep networks would also benefit from this technique, but leave it to future work.
 
@@
 \subsection{Final Comments on Stacked Autoencoders}
-As we mentioned in verification, if we used too few iterations for the network to converge, we saw a significant decrease in the performance.
+As we mentioned in section \ref{subsec:verification}, too few iterations led to a significant decrease in performance.
 Due to the complexity of the data, our final deep network has not fully converged and that is why its accuracy is so much worse than the baseline.

### 2013-12-18 postprocessors/clean_for_roc.py
Fix graph lables

@@ -87,8 +87,8 @@ plot(all_curve_x[1], all_curve_y[1])
 plot(all_curve_x[2], all_curve_y[2])
 legend(("ALPHA", "BETA", "LOOP"), loc='lower right')
 plot([0, .5, 1],[0, .5, 1], linestyle='--')
-xlabel('FPR')
-ylabel('TPR')
-title('ROC Curve for ' + desired_class + ' Classification')
-savefig(output_folder + 'ROC_all.png')
+xlabel('False Positive Rate')
+ylabel('True Positive Rate')
+title('ROC Curves for All Classifications')
+savefig(output_folder + 'ROC_all_final_deep_net_100.png')
 show()

### 2013-12-18 docs/paper.tex
paper done

@@
 We found that our best performance of 53.2\% resulted from parameters of window size 13, 6 hidden layers, and 26 units per hidden layer.
-Unfortunately, this underperforms as compared to other learning methods.
+Unfortunately, this under performs as compared to other learning methods.
 \end{abstract}
@@
 However, increasing the window size sharply increased the number of weights that needed to be trained.
-As a compromise between accuraccy and performance, we chose to use a window size of 13 for all other experiments.
+As a compromise between accuracy and performance, we chose to use a window size of 13 for all other experiments.
 
@@
 
-We found that these tests worked signifcantly worse than the standard deep networks.
+We found that these tests worked significantly worse than the standard deep networks.
 With three hidden layers and a decay factor of .25, we achieved an average accuracy of 50.37\%.
@@
 
-\section{Final Results and Discussion}
+\section{Results}
 % Determined optimal params - now use test set too
@@
 \subsection{Deep Networks}
+\label{subsec:deepnetresults}
 The final deep network has a window size of 13 and 6 hidden layers of size 26.
@@
 Despite the fact that the accuracy is lower than the baseline, all three ROC curves in figure \ref{fig:roc} show that the deep network's performance is better than simply guessing.
-Also note that for predictions with high confidence, our network predicts the secondary structure well, particularily beta classifications.
+Also note that for predictions with high confidence, our network predicts the secondary structure well, particularly beta classifications.
 
@@
 
-\subsection{Final Comments on Stacked Autoencoders}
-As we mentioned in section \ref{subsec:verification}, too few iterations led to a significant decrease in performance.
-Due to the complexity of the data, our final deep network has not fully converged and that is why its accuracy is so much worse than the baseline.
+\section{Final Comments on Stacked Autoencoders}
+As we mentioned in section \ref{subsec:verification}, too few iterations led to a significant decrease in accuracy.
+This is reinforced by the results seen in section \ref{subsec:deepnetresults}, showing that increasing the number of iterations improved our deep network's performance.
+Thus, we believe our final deep network has not fully converged, which is one reason why its accuracy is so much worse than the baseline.
 We are confident that with significantly longer training time (on the order of weeks), the deep network would fully converge.
@@
 
-\subsection{Lessons Learned}
+\section{Lessons Learned}
 In the process of working on this project, we learned that it is very difficult to understand the re-representations produced by the hidden layers of a large multi-layer network.
@@
 We also learned that deep networks need to be tested very carefully, since they have so many interconnections and weights.
-Overall, we feel that this was a valuable experience, gaining experience in neural networks.
+Overall, we feel that this study was a valuable experience.
 
@@
 We considered additional modifications to the basic algorithm that we were unable to test due to time constraints.
-We would have liked to experiment with additional functions for our hidden and output units, rather than only using a sigmoid.
-One interesting modification would be using a function that further encourages sparsity by using a penalty term \cite{ng}.
+We would have liked to experiment with additional activation functions for our hidden and output units, rather than only using a sigmoid.
+For example, it would be interesting to use a function that further encourages sparsity by using a penalty term \cite{Ng}.
 
 \section{Conclusion}
-We explored a deep network's ability to classsify secondary structures in proteins.
-We created a baseline of a standard single-hidden layer neural network for comparison.
-We did a parameter tuning study, and found that the best parameters for the deep network were a window size of 13 and 6 hidden layers with 26 units per layer, resulting in a final accuracy of 53.2\%.
+We explored a deep network's ability to classify secondary structures in proteins.
+We created a baseline of a standard single hidden layer neural network for comparison.
+We performed a parameter tuning study, and found that the best parameters for the deep network were a window size of 13 and 6 hidden layers with 26 units per layer, resulting in a final accuracy of 53.2\%.
 We found that our deep networks were unable to achieve the accuracy of a standard neural network.
@@
 G.E. Hinton and R. R. Salakhutdinov,
-   `` Reducing the dimensionality of data with neural networks"
+   ``Reducing the dimensionality of data with neural networks"
    in \emph{Science}, 2006.
@@
 Joosten RP, Te Beek TAH, Krieger E, Hekkelman ML, Hooft RWW, Schneider R, Sander C, Vriend G,
-   `` A series of PDB related databases for everyday needs"
+   ``A series of PDB related databases for everyday needs"
    in \emph{NAR}, 2010.
@@
 Kabsch W, Sander C,
-   `` Dictionary of protein secondary structure: pattern recognition of hydrogen-bonded and geometrical features"
+   ``Dictionary of protein secondary structure: pattern recognition of hydrogen-bonded and geometrical features"
    in \emph{Biopolymers}, 1983.
@@
 Andrew Ng,
-   `` CS294 Lecture notes: Sparse autoencoder"
+   ``CS294 Lecture notes: Sparse autoencoder"
    \textbf{www.stanford.edu/class/cs294a/}
@@
 Ning Qian and Terrence J. Sejnowski,
-  `` Predicting the secondary structure of globular proteins using neural network models"
+  ``Predicting the secondary structure of globular proteins using neural network models"
   in \emph{J. Mol. Bio.}, 1988.
@@
 Burkhard Rost and Christ Sander,
-  `` Prediction of protein secondary structures at bettter than 70\% accuracy"
+  ``Prediction of protein secondary structures at bettter than 70\% accuracy"
   in \emph{J. Mol. Bio.}, 1993.
@@
 Paul Stolorz, Alan Lapedes, and Yuan Xia,
-   `` Predicting protein secondary structure using neural net and statistical methods"
+   ``Predicting protein secondary structure using neural net and statistical methods"
    in \emph{J. Mol. Bio.}, 1992.
@@
    C.L. Blake and C.J. Merz,
-   `` UCI repository of machine learning databases"
+   ``UCI repository of machine learning databases"
    \textbf{http://archive.ics.uci.edu/ml/datasets/}
-   \textbf{Molecular+Biology+\}
+   \textbf{Molecular+Biology+}
    \textbf{(Protein+Secondary+Structure)}
